---
title: "Perennial Pepperweed Mapping"
subtitle: "Noxious Weeds Monitoring — LORP FY 2026–27 ([approved workplan PDF](https://inyowater.org/wp-content/uploads/2026/06/2026-2027-LORP-Work-Plan-and-Budget_draft_062226.pdf))"
tbl-cap-location: top   
affiliation: "Inyo County Water Department"
date: today
date-modified: today
citation:
  type: report
  container-title: "Data Report"
  publisher: "Inyo County Water Department"
  issued: today
  url: https://github.com/inyo-gov/noxious-weeds
google-scholar: true
params:
  current_year: 2025
  start_year: 2018
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 4
    toc-location: right
    code-fold: true
    code-tools: true
    css: styles.css
  docx:
    toc: true
    toc-depth: 3
    number-sections: true
    reference-doc: custom-reference.docx
execute:
  echo: false
  warning: false
  message: false
prefer-html: true
---

```{r setup}
#| message: false
#| warning: false
library(sf)
library(dplyr)
library(tidyr)
library(leaflet)
library(leaflet.extras2)
library(lubridate)
library(htmltools)
library(knitr)
library(DT)
library(ggplot2)
library(scales)
library(httr)

# Field abundance codes: 5, 15, 25, 100, 200 where 200 means ≥200 (often thousands)
format_abundance <- function(x) {
  ifelse(is.na(x), "Unknown",
         ifelse(x >= 200, "200+ (may be thousands)", as.character(x)))
}
```

## Summary

This data report covers the monitoring of **Lepidium latifolium** (perennial pepperweed) in the Lower Owens River Project (LORP) area from `r params$start_year` to `r params$current_year`. The data is sourced from the [ArcGIS Online Noxious Weeds 2025 feature service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0) and provides updated spatial occurrence as of `r params$current_year`. 

> "Perennial pepperweed (Lepidium latifolium), an introduced plant from southeastern Europe and Asia, is invasive throughout the western United States. It can establish in a wide range of environments and is a common problem in flood plains, irrigation structures, pastures, wetlands, riparian areas, roadsides, and residential site. Recent surveys identify perennial pepperweed as a weed problem in nearly all of California, and both the California Department of Food and Agriculture (CDFA) and California Invasive Plant Council (Cal-IPC) list it as a noxious weed of great ecological concern." - [UC IPM](https://ipm.ucanr.edu/home-and-landscape/perennial-pepperweed/#gsc.tab=0)

**California Invasive Plant Council (Cal-IPC) Rating: High**  
High – These species have severe ecological impacts on physical processes, plant and animal communities, and vegetation structure. Their reproductive biology and other attributes are conducive to moderate to high rates of dispersal and establishment. Most are widely distributed ecologically.

**California Department of Food and Agriculture (CDFA) Rating: -***  
* – included in the CCR Section 4500 list of California State Noxious Weeds.

```{r data-loading}
#| message: false
#| warning: false

# ArcGIS Feature Service URL
base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0/query?"

# Query parameters for all data
query_params <- paste0(
  "where=1=1",
  "&outFields=*",
  "&f=geojson",
  "&outSR=4326"
)

# Construct full query URL
query_url <- paste0(base_url, query_params)

# Load data from ArcGIS feature service
pepper_data <- st_read(query_url, quiet = TRUE)

# Process dates - convert Unix timestamps to dates
pepper_data <- pepper_data %>%
  mutate(
    Date_Observed = as.POSIXct(Date/1000, origin='1970-01-01'),
    Date_Created = as.POSIXct(CreationDate/1000, origin='1970-01-01'),
    Year_Observed = year(Date_Observed),
    Year_Created = year(Date_Created),
    # Use creation year as fallback when observation year is missing
    Year_Display = ifelse(is.na(Year_Observed), Year_Created, Year_Observed),
    Date_Display = as.POSIXct(ifelse(is.na(Date_Observed), as.numeric(Date_Created), as.numeric(Date_Observed)), origin='1970-01-01')
  )

# Filter to LORP area based on northernmost current year data point
pepper_current_year_for_bounds <- pepper_data %>%
  filter(Year_Display == params$current_year)

if(nrow(pepper_current_year_for_bounds) > 0) {
  # Get the northernmost latitude from current year data
  northernmost_lat <- max(st_coordinates(pepper_current_year_for_bounds)[,2])
  
  # Filter all data to LORP area (south of northernmost current year point)
  pepper_data <- pepper_data %>%
    filter(st_coordinates(.)[,2] <= northernmost_lat)
  
  cat("LORP Area Filter Applied:\n")
  cat("- Northernmost", params$current_year, "point latitude:", round(northernmost_lat, 4), "°N\n")
  cat("- Filtered dataset now contains", nrow(pepper_data), "observations\n")
} else {
  cat("No", params$current_year, "data found for LORP area filtering\n")
}

# Include all records (no year filtering)
# Separate data by display year (observation year with creation year fallback)
pepper_2018 <- pepper_data %>% filter(Year_Display == 2018)
pepper_2019 <- pepper_data %>% filter(Year_Display == 2019)
pepper_2020 <- pepper_data %>% filter(Year_Display == 2020)
pepper_2021 <- pepper_data %>% filter(Year_Display == 2021)
pepper_2022 <- pepper_data %>% filter(Year_Display == 2022)
pepper_2023 <- pepper_data %>% filter(Year_Display == 2023)
pepper_2024 <- pepper_data %>% filter(Year_Display == 2024)
pepper_2025 <- pepper_data %>% filter(Year_Display == 2025)

# Create recent vs historical comparison
pepper_recent <- pepper_data %>% filter(Year_Display >= 2022)
pepper_historical <- pepper_data %>% filter(Year_Display >= 2018 & Year_Display <= 2021)

# Summary statistics for all years (using display year with fallback)
summary_stats <- pepper_data %>%
  filter(!is.na(Year_Display)) %>%
  group_by(Year_Display) %>%
  summarise(
    Total_Observations = n(),
    Avg_Abundance = round(mean(Abundance, na.rm = TRUE), 1),
    Avg_Height = round(mean(Height, na.rm = TRUE), 1),
    Max_Abundance = max(Abundance, na.rm = TRUE),
    Records_With_Obs_Date = sum(!is.na(Year_Observed)),
    Records_With_Creation_Date_Only = sum(is.na(Year_Observed) & !is.na(Year_Created)),
    .groups = 'drop'
  ) %>%
  arrange(Year_Display)
```

### Data Overview

The dataset includes `r nrow(pepper_data)` total observations across `r length(unique(pepper_data$Year_Display[!is.na(pepper_data$Year_Display)]))` years, with peak survey activity in `r names(sort(table(pepper_data$Year_Display, useNA='ifany'), decreasing=TRUE)[1])` (`r max(table(pepper_data$Year_Display, useNA='ifany'))` records).

### Key Findings Summary

- **Total Observations**: `r nrow(pepper_data)` records across all years (2018-2025)
- **Peak Survey Year**: `r names(sort(table(pepper_data$Year_Display, useNA='ifany'), decreasing=TRUE)[1])` with `r max(table(pepper_data$Year_Display, useNA='ifany'))` observations
- **Recent Activity**: `r nrow(pepper_recent)` observations in 2022-2025
- **Historical Baseline**: `r nrow(pepper_historical)` observations in 2018-2021
- **Data Coverage**: `r round((nrow(pepper_data) - sum(is.na(pepper_data$Year_Display))) / nrow(pepper_data) * 100, 1)`% of records have complete date information (using observation date with creation date fallback)

## Interactive Map Comparison

The following interactive map shows pepperweed populations across all years (`r params$start_year`-`r params$current_year`). Use the layer controls in the top-right corner to toggle between different years. The map includes all `r nrow(pepper_data)` documented observations, with feature creation dates used when observation dates are unavailable.

**Map layers:** purple **River channels** (multithreaded Owens River) and blue **High-water line** (June 2023 floodplain). Survey-gap polygons are deferred for now — use the tabular monitoring recommendations below.

::: {.map-a11y-summary}
**Accessible map summary:** Points are colored by survey year (`r params$start_year`–`r params$current_year`) along the LORP corridor. Toggle year layers and basemap with the map controls. The same occurrence data are available as downloadable GeoJSON/KML below and are summarized in the reach, occupancy, and river-mile tables later in this report. Screen-reader users can rely on those tables and the year legend title “Detected and reported per year” rather than the interactive map canvas.
:::

<div class="download-section">
<h3>Data Downloads</h3>
<div class="download-links">
<a href="pepperweed_data.geojson">Download GeoJSON</a>
<a href="pepperweed_data.kml">Download KML</a>
</div>
<p>Complete dataset with all `r nrow(pepper_data)` pepperweed observations</p>
</div>

```{r generate-geojson}
#| message: false
#| warning: false
#| echo: false
#| results: hide

# Generate GeoJSON file for download
st_write(pepper_data, "pepperweed_data.geojson", driver = "GeoJSON", delete_dsn = TRUE, quiet = TRUE)

# Generate KML file for Google Earth (will be zipped to KMZ)
st_write(pepper_data, "pepperweed_data.kml", driver = "KML", delete_dsn = TRUE, quiet = TRUE)

# Calculate spatial extent
bbox <- st_bbox(pepper_data)
extent_info <- list(
  north = bbox$ymax,
  south = bbox$ymin, 
  east = bbox$xmax,
  west = bbox$xmin,
  center_lat = (bbox$ymax + bbox$ymin) / 2,
  center_lon = (bbox$xmax + bbox$xmin) / 2
)
```

<div class="spatial-extent">
<h4>Study Area Boundaries</h4>
<ul>
<li><strong>North</strong>: `r round(extent_info$north, 4)`°N</li>
<li><strong>South</strong>: `r round(extent_info$south, 4)`°N</li>
<li><strong>East</strong>: `r round(extent_info$east, 4)`°W</li>
<li><strong>West</strong>: `r round(extent_info$west, 4)`°W</li>
<li><strong>Center</strong>: `r round(extent_info$center_lat, 4)`°N, `r round(extent_info$center_lon, 4)`°W</li>
</ul>
</div>

```{r interactive-map}
#| message: false
#| warning: false

# Add reach information to pepper_data if available
if(exists("pepper_data_with_reaches")) {
  # Merge reach information back to main dataset
  pepper_data <- pepper_data %>%
    left_join(
      pepper_data_with_reaches %>% 
        st_drop_geometry() %>% 
        select(OBJECTID, Name),
      by = "OBJECTID"
    )
}

# Load water quality stations as reference points
wq_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/lorp_wq_view/FeatureServer/0/query?where=1=1&outFields=*&f=geojson&outSR=4326"
wq_stations <- st_read(wq_url, quiet = TRUE)

# Load river reaches for labeling (if not already loaded)
if(!exists("river_reaches")) {
  reach_url <- "https://inyocounty.maps.arcgis.com/sharing/rest/content/items/90e5870bd5914a928bd97b023f07b807/data"
  river_reaches <- st_read(reach_url, quiet = TRUE)
  river_reaches <- st_transform(river_reaches, st_crs(pepper_data))
}

# Get reach centroids for labeling - only LORP reaches 1-6
reach_centroids <- river_reaches %>%
  filter(Name %in% c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6")) %>%
  st_centroid() %>%
  mutate(Reach_Label = gsub("LORP Reach ", "R", Name))

# High-water floodplain + multithreaded river (reference layers)
fp_geojson <- "exports/survey_nav_2026/lorp_highwater_floodplain.geojson"
ch_geojson <- "exports/survey_nav_2026/owens_river_channels.geojson"
if ((!file.exists(fp_geojson) || !file.exists(ch_geojson)) && file.exists("export_survey_nav.R")) {
  system2("Rscript", "export_survey_nav.R", stdout = FALSE, stderr = FALSE)
}
floodplain <- NULL
river_channels <- NULL
if (file.exists(fp_geojson)) {
  floodplain <- st_read(fp_geojson, quiet = TRUE) %>% st_transform(4326)
}
if (file.exists(ch_geojson)) {
  river_channels <- st_read(ch_geojson, quiet = TRUE) %>% st_transform(4326)
}

# Create comprehensive map with all years
combined_map <- leaflet(pepper_data) %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery") %>%
  addProviderTiles(providers$OpenStreetMap, group = "OpenStreetMap")

# June 2023 high-water floodplain (full extent)
if (!is.null(floodplain) && nrow(floodplain) > 0) {
  combined_map <- combined_map %>%
    addPolygons(
      data = floodplain,
      fillColor = "#64b5f6",
      fillOpacity = 0.18,
      color = "#0d47a1",
      weight = 2,
      opacity = 0.9,
      popup = ~paste0("<b>June 2023 high-water line</b><br>", Acres, " acres"),
      group = "High-water line",
      options = pathOptions(clickable = TRUE)
    )
}

# Multithreaded Owens River channels
if (!is.null(river_channels) && nrow(river_channels) > 0) {
  combined_map <- combined_map %>%
    addPolylines(
      data = river_channels,
      color = "#4527a0",
      weight = 3,
      opacity = 0.95,
      popup = ~paste0("<b>Owens River channel</b><br>", ifelse(is.na(NAME), "", NAME)),
      group = "River channels"
    )
}

combined_map <- combined_map %>%
  # Add 2018 markers
  addCircleMarkers(
    data = pepper_2018,
    radius = 5,
    color = "black",
    fillColor = "black",
    fillOpacity = 0.7,
    weight = 2,
      popup = ~paste0(
        "<b>2018 Data</b><br>",
        "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
        "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
        "Species: ", Species, "<br>",
        "Abundance: ", format_abundance(Abundance), "<br>",
        "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
        ifelse(exists("Name") && !is.na(Name), paste0("LORP Reach: ", Name, "<br>"), ""),
        "Notes: ", ifelse(is.na(Notes), "None", Notes)
      ),
    group = "2018 Data"
  ) %>%
  # Add 2019 markers
  addCircleMarkers(
    data = pepper_2019,
    radius = 5,
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2019 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2019 Data"
  ) %>%
  # Add 2020 markers
  addCircleMarkers(
    data = pepper_2020,
    radius = 5,
    color = "red",
    fillColor = "red",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2020 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2020 Data"
  ) %>%
  # Add 2021 markers
  addCircleMarkers(
    data = pepper_2021,
    radius = 5,
    color = "yellow",
    fillColor = "yellow",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2021 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2021 Data"
  ) %>%
  # Add 2022 markers
  addCircleMarkers(
    data = pepper_2022,
    radius = 5,
    color = "orange",
    fillColor = "orange",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2022 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2022 Data"
  ) %>%
  # Add 2023 markers
  addCircleMarkers(
    data = pepper_2023,
    radius = 5,
    color = "green",
    fillColor = "green",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2023 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2023 Data"
  ) %>%
  # Add 2024 markers
  addCircleMarkers(
    data = pepper_2024,
    radius = 5,
    color = "blue",
    fillColor = "blue",
    fillOpacity = 0.8,
    weight = 3,
    popup = ~paste0(
      "<b>2024 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2024 Data"
  ) %>%
  # Add 2025 markers
  addCircleMarkers(
    data = pepper_2025,
    radius = 5,
    color = "purple",
    fillColor = "purple",
    fillOpacity = 0.9,
    weight = 3,
    popup = ~paste0(
      "<b>2025 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", format_abundance(Abundance), "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      ifelse(exists("Name") && !is.na(Name), paste0("LORP Reach: ", Name, "<br>"), ""),
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2025 Data"
  ) %>%
  # Add water quality stations as reference points (if data available)
  {if(nrow(wq_stations) > 0) {
    addCircleMarkers(
      data = wq_stations,
      radius = 4,
      color = "white",
      fillColor = "blue",
      fillOpacity = 0.8,
      weight = 2,
      popup = ~paste0("<b>Water Quality Station</b><br>Reference Point"),
      group = "Water Quality Stations"
    )
  } else {
    .  # No stations to add
  }} %>%
  # Add reach labels (centroids)
  addLabelOnlyMarkers(
    data = reach_centroids,
    label = ~Reach_Label,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = "center",
      textOnly = TRUE,
      style = list(
        "color" = "white",
        "font-size" = "14px",
        "font-weight" = "bold",
        "text-shadow" = "1px 1px 2px rgba(0,0,0,0.8)"
      )
    ),
    group = "Reach Labels"
  )

overlay_groups <- c(
  if (!is.null(floodplain) && nrow(floodplain) > 0) "High-water line",
  if (!is.null(river_channels) && nrow(river_channels) > 0) "River channels",
  "2018 Data", "2019 Data", "2020 Data", "2021 Data", "2022 Data",
  "2023 Data", "2024 Data", "2025 Data", "Reach Labels",
  if (nrow(wq_stations) > 0) "Water Quality Stations"
)

combined_map <- combined_map %>%
  addLayersControl(
    baseGroups = c("World Imagery", "OpenStreetMap"),
    overlayGroups = overlay_groups,
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomleft",
    colors = c("black", "darkred", "red", "yellow", "orange", "green", "blue", "purple", if(nrow(wq_stations) > 0) "blue"),
    labels = c("2018", "2019", "2020", "2021", "2022", "2023", "2024", "2025", if(nrow(wq_stations) > 0) "WQ Stations"),
    title = "Detected and reported per year"
  )

combined_map <- combined_map %>%
  htmlwidgets::onRender(
    "function(el, x) {
      el.setAttribute('role', 'img');
      el.setAttribute(
        'aria-label',
        'Interactive map of perennial pepperweed detections along the Lower Owens River Project corridor, colored by survey year. Use the layer control to toggle years. Equivalent data are in the downloadable files and tables in this report.'
      );
    }"
  )

combined_map
```

### `r params$current_year` Distribution - plants per site estimated from field observations

Field abundance is recorded in discrete estimate classes (**5, 15, 25, 100, 200**). A value of **200 means ≥200 plants** — a lower bound for dense stands that may contain hundreds to thousands of individuals (not an exact count of 200).

The following chart shows the distribution of these abundance classes for pepperweed sites documented in `r params$current_year`:

```{r abundance-histogram}
#| message: false
#| warning: false
#| fig-alt: "Bar chart of perennial pepperweed sites by field abundance class (5, 15, 25, 100, 200+) for the current survey year. Class 200+ means 200 or more plants and may represent thousands."

# Filter for current year data only
pepper_current_year <- pepper_data %>%
  filter(Year_Display == params$current_year)

# Create abundance histogram
if(nrow(pepper_current_year) > 0) {
  # Use field estimate classes (200 = 200+)
  abundance_data <- pepper_current_year %>%
    filter(!is.na(Abundance)) %>%
    mutate(
      Abundance_Bin = case_when(
        Abundance <= 5 ~ "5",
        Abundance <= 15 ~ "15",
        Abundance <= 25 ~ "25",
        Abundance <= 100 ~ "100",
        TRUE ~ "200+"
      ),
      Abundance_Bin = factor(Abundance_Bin,
                            levels = c("5", "15", "25", "100", "200+"),
                            ordered = TRUE)
    )
  
  # Only include bins that have data
  abundance_counts <- abundance_data %>%
    count(Abundance_Bin) %>%
    filter(n > 0)
  
  # Create bar chart with only populated bins
  hist_plot <- ggplot(abundance_counts, aes(x = Abundance_Bin, y = n)) +
    geom_bar(stat = "identity", fill = "#2E86AB", color = "white", alpha = 0.8, width = 0.7) +
    labs(
      title = paste("Distribution of Pepperweed Abundance in", params$current_year),
      subtitle = paste("Total sites:", nrow(pepper_current_year),
                       "· Class 200+ = ≥200 plants (may be thousands)"),
      x = "Abundance class (field estimate)",
      y = "Number of Sites"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 12, color = "#666666"),
      axis.title = element_text(size = 12),
      axis.text.x = element_text(angle = 0, hjust = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank()
    ) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 8))
  
  print(hist_plot)
  
  # Management categories
  small_pops <- sum(pepper_current_year$Abundance < 200, na.rm = TRUE)
  large_pops <- sum(pepper_current_year$Abundance >= 200, na.rm = TRUE)
  
  cat("###", params$current_year, "Site Distribution\n")
  cat("- **Total Sites**:", nrow(pepper_current_year), "\n")
  cat("- **Sites below 200 class**:", small_pops, "(", round(small_pops/nrow(pepper_current_year)*100, 1), "%)\n")
  cat("- **Sites in 200+ class (≥200 plants; may be thousands)**:", large_pops, "(", round(large_pops/nrow(pepper_current_year)*100, 1), "%)\n")
  
  # Detailed bin counts
  cat("\n### Detailed Abundance Distribution\n")
  for(i in 1:nrow(abundance_counts)) {
    bin_name <- as.character(abundance_counts$Abundance_Bin[i])
    bin_count <- abundance_counts$n[i]
    label <- if (bin_name == "200+") "200+ plants (≥200; may be thousands)" else paste0(bin_name, " plants (estimate class)")
    cat("- **", label, "**: ", bin_count, " sites\n", sep = "")
  }
  
  # Range information
  cat("\n### Abundance coding\n")
  cat("- Field classes used: **5, 15, 25, 100, 200**\n")
  cat("- **200 = 200 or more** — dense patches are not counted stem-by-stem and can represent thousands of individuals\n")
} else {
  cat("No", params$current_year, "data available for abundance analysis.\n")
}
```

The interactive map above provides the primary visualization of pepperweed distribution patterns. 


#### Recent Sites
```{r new-sites}
#| message: false
#| warning: false

# Show all recent sites (2024-2025) - prioritize small sites for early intervention
# Sort by numeric abundance, then format for display (200 = 200+)
recent_sites <- pepper_data %>%
  filter(Year_Display >= 2024) %>%
  arrange(Abundance, desc(Year_Display)) %>%
  mutate(Abundance = format_abundance(Abundance)) %>%
  select(Date_Display, Abundance, Height, Notes, Year_Display)

if(nrow(recent_sites) > 0) {
  datatable(recent_sites,
            caption = "Recent Pepperweed Sites (2024-2025) - Prioritize Small Sites for Early Intervention",
            options = list(pageLength = 15, scrollX = TRUE))
} else {
  cat("No recent sites found.")
}
```




## Data Quality and Limitations

### Data Sources
- **Primary Source**: [ArcGIS Online Noxious Weeds 2025 Feature Service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0)

- **Last Updated**: `r format(as.POSIXct(max(pepper_data$CreationDate, na.rm = TRUE)/1000, origin='1970-01-01'), "%B %d, %Y")`
- **Coordinate System**: WGS84 (EPSG:4326)

### Limitations
- Data represents documented observations only
- Population estimates may vary based on survey timing and conditions
- Some areas may be under-sampled due to access constraints
- Abundance is recorded in field estimate classes (**5, 15, 25, 100, 200**); **200 means ≥200 plants** and dense stands may contain thousands of individuals

## River Reach Analysis

```{r river-reach-analysis}
#| message: false
#| warning: false
#| fig-width: 16
#| fig-height: 15
#| fig-alt: "Faceted bar charts showing pepperweed site counts by year for each LORP reach (Reaches 1 through 6)."

# Read river reach polygons from ArcGIS Online and filter to LORP reaches 1-6
reach_url <- "https://inyocounty.maps.arcgis.com/sharing/rest/content/items/90e5870bd5914a928bd97b023f07b807/data"
river_reaches <- st_read(reach_url, quiet = TRUE)

# Filter to only LORP reaches 1-6 - explicit selection
desired_lorp_reaches <- c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6")
river_reaches <- river_reaches %>%
  filter(Name %in% desired_lorp_reaches)

# Ensure both datasets have the same CRS (WGS84)
river_reaches <- st_transform(river_reaches, st_crs(pepper_data))

# Check if we have reach data
cat("Number of river reaches loaded:", nrow(river_reaches), "\n")
cat("Reach names:", unique(river_reaches$Name), "\n")

# Perform spatial join to determine which reach each pepperweed site is closest to
# Use nearest neighbor join since sites might not fall exactly within reach polygons
pepper_data_with_reaches <- pepper_data %>%
  st_join(river_reaches, join = st_nearest_feature)

# Check how many sites were matched
cat("Sites before filtering:", nrow(pepper_data), "\n")
cat("Sites after spatial join:", nrow(pepper_data_with_reaches), "\n")
cat("Sites with reach assignments:", sum(!is.na(pepper_data_with_reaches$Name)), "\n")

# Filter to only sites within reaches
pepper_data_with_reaches <- pepper_data_with_reaches %>%
  filter(!is.na(Name))  # Only keep sites that fall within a reach

# Create summary data for reach analysis
reach_summary <- pepper_data_with_reaches %>%
  st_drop_geometry() %>%
  group_by(Year_Display, Name) %>%
  summarise(Site_Count = n(), .groups = 'drop') %>%
  # Ensure all year-reach combinations are represented
  complete(Year_Display = unique(Year_Display), 
           Name = unique(Name), 
           fill = list(Site_Count = 0))

# Create faceted histogram for reaches
if(nrow(reach_summary) > 0) {
  # Cap sites at 20 for better small site visibility
  reach_summary_capped <- reach_summary %>%
    mutate(Site_Count_Capped = pmin(Site_Count, 20))
  
  # Create ordered factor for reaches (1 at top, 6 at bottom)
  reach_summary_capped <- reach_summary_capped %>%
    mutate(Name_Ordered = factor(Name, levels = rev(c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6"))))
  
  reach_plot <- ggplot(reach_summary_capped, aes(x = Site_Count_Capped, y = Name_Ordered)) +
    geom_col(fill = "red", alpha = 0.8, width = 0.7) +
    facet_wrap(~ Year_Display, ncol = 8, scales = "fixed") +  # Single row, fixed scales
    scale_x_continuous(limits = c(0, 20), breaks = c(0, 5, 10, 15, 20)) +  # Cap at 20
    scale_y_discrete() +  # Reach names
    labs(
      title = "Pepperweed Sites by LORP Reach and Year",
      subtitle = "Distribution of sites within LORP river reaches (capped at 20 sites)",
      x = "Number of Sites (capped at 20)",
      y = "LORP Reach"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 18, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 16, color = "#666666"),
      axis.title = element_text(size = 14),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      strip.text = element_text(size = 13, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(25, 25, 25, 25)
    )
  
  print(reach_plot)
  
  # Summary statistics
  cat("### LORP Reach Distribution Summary\n")
  cat("**Total Reaches Analyzed**:", length(unique(reach_summary$Name)), "\n")
  cat("**Years Covered**:", min(reach_summary$Year_Display), "-", max(reach_summary$Year_Display), "\n")
  cat("**Reaches with Data**:", length(unique(reach_summary$Name[reach_summary$Site_Count > 0])), "\n")
  
  # Most active reaches
  top_reaches <- reach_summary %>%
    group_by(Name) %>%
    summarise(Total_Sites = sum(Site_Count), .groups = 'drop') %>%
    arrange(desc(Total_Sites))
  
  cat("\n### Reach Activity Summary\n")
  for(i in 1:nrow(top_reaches)) {
    cat("- **", top_reaches$Name[i], "**:", top_reaches$Total_Sites[i], "sites\n")
  }
  
  # Sites not in any reach
  sites_outside_reaches <- nrow(pepper_data) - nrow(pepper_data_with_reaches)
  if(sites_outside_reaches > 0) {
    cat("\n### Sites Outside LORP Reaches\n")
    cat("- **Sites outside reach polygons**:", sites_outside_reaches, "\n")
    cat("- **Percentage of total sites**:", round(sites_outside_reaches/nrow(pepper_data)*100, 1), "%\n")
  }
} else {
  cat("No reach data available for analysis.\n")
}
```

## Occupancy Analysis
```{r occupancy-analysis}
#| message: false
#| warning: false

# Load river miles data if not already loaded
if(!exists("rivermiles")) {
  rivermiles <- st_read('data/LORP_RiverMiles_revised.shp', quiet = TRUE)
  rivermiles <- rivermiles %>%
    mutate(RiverMile = as.numeric(NAME))
  rivermiles <- st_transform(rivermiles, st_crs(pepper_data))
}

# Create 1/10 mile segments for detailed occupancy analysis
# Total of 600 segments across 60 miles (0.1 mile resolution)
segment_size <- 0.1
total_miles <- 60
total_segments <- total_miles / segment_size  # 600 segments

# Create segment boundaries
segment_boundaries <- data.frame(
  Segment_ID = 1:total_segments,
  Mile_Start = seq(0, total_miles - segment_size, by = segment_size),
  Mile_End = seq(segment_size, total_miles, by = segment_size)
)

# Join pepperweed data with river miles and reaches
pepper_data_with_miles_and_reaches <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

# Determine which segment each pepperweed site falls into
pepper_data_with_segments <- pepper_data_with_miles_and_reaches %>%
  st_drop_geometry() %>%
  mutate(
    Segment_ID = floor(RiverMile / segment_size) + 1,
    # Ensure segments are within bounds
    Segment_ID = pmin(Segment_ID, total_segments)
  )

# Overall occupancy by year
overall_occupancy <- pepper_data_with_segments %>%
  group_by(Year_Display) %>%
  summarise(
    Segments_Occupied = n_distinct(Segment_ID),
    Total_Segments = total_segments,
    Occupancy_Percent = round(Segments_Occupied / Total_Segments * 100, 1),
    .groups = 'drop'
  ) %>%
  arrange(Year_Display)

# Reach-specific occupancy
reach_occupancy <- pepper_data_with_segments %>%
  group_by(Year_Display, Name) %>%
  summarise(
    Segments_Occupied = n_distinct(Segment_ID),
    .groups = 'drop'
  ) %>%
  # Calculate total segments per reach
  left_join(
    pepper_data_with_miles_and_reaches %>%
      st_drop_geometry() %>%
      group_by(Name) %>%
      summarise(
        Reach_Mile_Start = min(RiverMile, na.rm = TRUE),
        Reach_Mile_End = max(RiverMile, na.rm = TRUE),
        Reach_Total_Segments = (Reach_Mile_End - Reach_Mile_Start) / segment_size,
        .groups = 'drop'
      ),
    by = "Name"
  ) %>%
  mutate(
    Occupancy_Fraction = round(Segments_Occupied / Reach_Total_Segments, 3),
    Occupancy_Percent = round(Occupancy_Fraction * 100, 1)
  ) %>%
  arrange(Year_Display, Name)

# Display results
cat("### Overall Occupancy Analysis\n")
cat("**Resolution**: 1/10 mile segments (600 total segments across 60 miles)\n\n")

# Overall occupancy table
cat("#### Overall Occupancy by Year\n")
print(kable(overall_occupancy, 
            col.names = c("Year", "Segments Occupied", "Total Segments", "Occupancy %"),
            align = c("c", "c", "c", "c"),
            caption = "Number of 1/10 mile segments occupied by pepperweed"))

# Reach occupancy table
cat("\n#### Reach-Specific Occupancy\n")
reach_occupancy_wide <- reach_occupancy %>%
  select(Year_Display, Name, Occupancy_Percent) %>%
  pivot_wider(names_from = Year_Display, values_from = Occupancy_Percent, names_prefix = "Year_") %>%
  arrange(Name)

print(kable(reach_occupancy_wide, 
            col.names = c("LORP Reach", "2018", "2019", "2020", "2021", "2022", "2023", "2024", "2025"),
            align = c("l", rep("c", 8)),
            caption = "Percentage of each reach occupied by pepperweed (by year)"))

# Summary statistics
cat("\n### Occupancy Summary Statistics\n")
cat("**Average Annual Occupancy**:", round(mean(overall_occupancy$Occupancy_Percent), 1), "%\n")
cat("**Highest Occupancy Year**:", overall_occupancy$Year_Display[which.max(overall_occupancy$Occupancy_Percent)], 
    "(", max(overall_occupancy$Occupancy_Percent), "%)\n")
cat("**Lowest Occupancy Year**:", overall_occupancy$Year_Display[which.min(overall_occupancy$Occupancy_Percent)], 
    "(", min(overall_occupancy$Occupancy_Percent), "%)\n")

# Most/least occupied reaches
reach_summary <- reach_occupancy %>%
  group_by(Name) %>%
  summarise(
    Avg_Occupancy = round(mean(Occupancy_Percent), 1),
    Max_Occupancy = max(Occupancy_Percent),
    .groups = 'drop'
  ) %>%
  arrange(desc(Avg_Occupancy))

cat("\n**Most Occupied Reach**:", reach_summary$Name[1], "(", reach_summary$Avg_Occupancy[1], "% average)\n")
cat("**Least Occupied Reach**:", reach_summary$Name[nrow(reach_summary)], "(", reach_summary$Avg_Occupancy[nrow(reach_summary)], "% average)\n")
```

### Cumulative Discovery and Occupancy Forecasts

Annual occupancy fluctuates with survey effort. A more informative metric for the “more you look, the more you find” pattern is **cumulative discovery**: the share of corridor segments that have ever had a detection. Pepperweed is already widespread on adjacent conveyances and upper-river reaches, so the planning question is when known occupancy along the LORP corridor approaches saturation.

#### Cumulative Discovery by Year

Once a segment or river mile has a detection, it remains in the cumulative total. This isolates discovery/spread from year-to-year survey intensity.

```{r occupancy-forecast}
#| message: false
#| warning: false
#| results: 'asis'

# Cumulative discovery: first year each 0.1-mi segment was detected
first_detect_seg <- pepper_data_with_segments %>%
  group_by(Segment_ID) %>%
  summarise(FirstYear = min(Year_Display), .groups = "drop")

# Cumulative discovery at whole-mile resolution
first_detect_mile <- pepper_data_with_segments %>%
  mutate(Mile = floor(RiverMile)) %>%
  group_by(Mile) %>%
  summarise(FirstYear = min(Year_Display), .groups = "drop")

obs_years <- sort(unique(pepper_data_with_segments$Year_Display))
n_miles_corridor <- length(0:60)  # 61 whole miles

cum_occupancy <- data.frame(
  Year = obs_years,
  Segments_Known = sapply(obs_years, function(y) sum(first_detect_seg$FirstYear <= y)),
  Miles_Known = sapply(obs_years, function(y) sum(first_detect_mile$FirstYear <= y))
) %>%
  mutate(
    Seg_Pct = round(Segments_Known / total_segments * 100, 1),
    Mile_Pct = round(Miles_Known / n_miles_corridor * 100, 1),
    New_Segments = Segments_Known - lag(Segments_Known, default = 0),
    New_Miles = Miles_Known - lag(Miles_Known, default = 0)
  )

print(kable(cum_occupancy %>% select(Year, Segments_Known, New_Segments, Seg_Pct, Miles_Known, New_Miles, Mile_Pct),
            col.names = c("Year", "Segments Known", "New Segments", "Segment %", "Miles Known", "New Miles", "Mile %"),
            align = c("c", "c", "c", "c", "c", "c", "c"),
            caption = "Cumulative pepperweed discovery along the LORP corridor (0.1-mi segments and whole river miles)"))

# --- Forecast helpers ---
# Linear: p = a + b * year  (extrapolate to 100)
# Asymptotic (negative exponential to K=100): p = K * (1 - exp(-r * (year - t0)))
# Logistic to K=100: p = K / (1 + exp(-r * (year - t_mid)))

fit_linear_to_100 <- function(years, pct, label) {
  fit <- lm(pct ~ years)
  a <- coef(fit)[1]
  b <- coef(fit)[2]
  year_100 <- if (b > 0) (100 - a) / b else Inf
  list(label = label, model = "Linear", a = a, b = b, year_100 = year_100, fit = fit,
       predict = function(y) pmin(100, pmax(0, a + b * y)))
}

fit_asymptotic_to_100 <- function(years, pct, label, K = 100) {
  # p = K * (1 - exp(-r * (t - t0))); constrain t0 near first year, fit r
  obj <- function(par) {
    r <- exp(par[1])  # > 0
    t0 <- par[2]
    pred <- K * (1 - exp(-r * (years - t0)))
    sum((pct - pred)^2)
  }
  starts <- list(
    c(log(0.05), min(years) - 1),
    c(log(0.08), min(years) - 2),
    c(log(0.03), min(years)),
    c(log(0.1), min(years) - 3)
  )
  best <- NULL
  best_sse <- Inf
  for (s in starts) {
    opt <- tryCatch(
      optim(s, obj, method = "L-BFGS-B",
            lower = c(log(1e-4), min(years) - 20),
            upper = c(log(1), max(years))),
      error = function(e) NULL
    )
    if (!is.null(opt) && opt$value < best_sse) {
      best_sse <- opt$value
      best <- opt
    }
  }
  r <- exp(best$par[1])
  t0 <- best$par[2]
  # year when p reaches target: t = t0 - log(1 - target/K) / r
  year_at <- function(target) t0 - log(1 - target / K) / r
  list(label = label, model = "Asymptotic (→100%)", r = r, t0 = t0, year_100 = Inf,
       year_99 = year_at(99), year_95 = year_at(95), year_90 = year_at(90),
       predict = function(y) K * (1 - exp(-r * (y - t0))))
}

fit_logistic_to_100 <- function(years, pct, label, K = 100) {
  # p = K / (1 + exp(-r * (t - t_mid)))
  obj <- function(par) {
    r <- exp(par[1])
    t_mid <- par[2]
    pred <- K / (1 + exp(-r * (years - t_mid)))
    sum((pct - pred)^2)
  }
  # Multiple starts — early-phase series are sensitive to initial midpoint
  starts <- list(
    c(log(0.2), mean(years)),
    c(log(0.15), max(years) + 3),
    c(log(0.25), max(years) + 5),
    c(log(0.1), max(years) + 10),
    c(log(0.3), max(years))
  )
  best <- NULL
  best_sse <- Inf
  for (s in starts) {
    opt <- tryCatch(
      optim(s, obj, method = "L-BFGS-B",
            lower = c(log(1e-4), min(years) - 5),
            upper = c(log(1), max(years) + 80)),
      error = function(e) NULL
    )
    if (!is.null(opt) && opt$value < best_sse) {
      best_sse <- opt$value
      best <- opt
    }
  }
  r <- exp(best$par[1])
  t_mid <- best$par[2]
  year_at <- function(target) t_mid - log(K / target - 1) / r
  list(label = label, model = "Logistic (K=100%)", r = r, t_mid = t_mid, sse = best_sse,
       year_50 = year_at(50), year_90 = year_at(90), year_95 = year_at(95),
       year_99 = year_at(99), year_100 = Inf,
       predict = function(y) K / (1 + exp(-r * (y - t_mid))))
}

# Fit both resolutions
seg_lin <- fit_linear_to_100(cum_occupancy$Year, cum_occupancy$Seg_Pct, "0.1-mi segments")
seg_asy <- fit_asymptotic_to_100(cum_occupancy$Year, cum_occupancy$Seg_Pct, "0.1-mi segments")
seg_log <- fit_logistic_to_100(cum_occupancy$Year, cum_occupancy$Seg_Pct, "0.1-mi segments")

mile_lin <- fit_linear_to_100(cum_occupancy$Year, cum_occupancy$Mile_Pct, "Whole river miles")
mile_asy <- fit_asymptotic_to_100(cum_occupancy$Year, cum_occupancy$Mile_Pct, "Whole river miles")
mile_log <- fit_logistic_to_100(cum_occupancy$Year, cum_occupancy$Mile_Pct, "Whole river miles")

# Horizon years for plot
last_obs <- max(cum_occupancy$Year)
# Recent-trajectory scenario: whole-mile slope over last ~2 years
# (hypothesis: limited annual survey understates ubiquity; recent discovery rate continues)
recent_miles <- cum_occupancy %>% filter(Year >= last_obs - 2)
mile_recent_fit <- lm(Mile_Pct ~ Year, data = recent_miles)
mile_recent_rate <- coef(mile_recent_fit)[2]
mile_recent_a <- coef(mile_recent_fit)[1]
mile_recent_year_100 <- if (mile_recent_rate > 0) (100 - mile_recent_a) / mile_recent_rate else Inf
mile_recent_predict <- function(y) pmin(100, pmax(0, mile_recent_a + mile_recent_rate * y))

# Extend far enough to show approach to 100% for mile-scale logistic/linear
horizon_end <- max(
  ceiling(mile_lin$year_100) + 2,
  ceiling(seg_lin$year_100) + 2,
  ceiling(mile_log$year_99) + 5,
  ceiling(seg_log$year_99) + 5,
  ceiling(mile_recent_year_100) + 2,
  last_obs + 25
)
horizon_end <- min(horizon_end, last_obs + 80)  # cap plot range
forecast_years <- seq(min(cum_occupancy$Year), horizon_end, by = 1)

# Build forecast table of milestone years
milestone_row <- function(fit_lin, fit_asy, fit_log, resolution) {
  year_lin <- function(target) {
    y <- (target - fit_lin$a) / fit_lin$b
    if (!is.finite(y) || fit_lin$b <= 0) return(NA_real_)
    y
  }
  data.frame(
    Resolution = resolution,
    Metric = c("50%", "75%", "90%", "95%", "99%", "100% (linear only)"),
    Linear = round(c(year_lin(50), year_lin(75), year_lin(90), year_lin(95), year_lin(99), fit_lin$year_100), 1),
    Asymptotic = round(c(
      fit_asy$t0 - log(1 - 50/100) / fit_asy$r,
      fit_asy$t0 - log(1 - 75/100) / fit_asy$r,
      fit_asy$year_90,
      fit_asy$year_95,
      fit_asy$year_99,
      NA_real_
    ), 1),
    Logistic = round(c(
      fit_log$year_50,
      fit_log$t_mid - log(100/75 - 1) / fit_log$r,
      fit_log$year_90,
      fit_log$year_95,
      fit_log$year_99,
      NA_real_
    ), 1)
  )
}

milestones <- bind_rows(
  milestone_row(seg_lin, seg_asy, seg_log, "0.1-mi segments"),
  milestone_row(mile_lin, mile_asy, mile_log, "Whole river miles")
)

# Horizontal lag: years until segment occupancy catches a given mile occupancy level
# (time to "infill" every 0.1-mi subsegment once whole miles reach that level)
year_at_logistic <- function(fit, target, K = 100) {
  if (target >= K) return(Inf)
  if (target <= 0) return(-Inf)
  fit$t_mid - log(K / target - 1) / fit$r
}

infill_levels <- c(50, 75, 90)
infill_gaps <- data.frame(
  Occupancy = paste0(infill_levels, "%"),
  Year_WholeMile = round(vapply(infill_levels, function(p) year_at_logistic(mile_log, p), numeric(1)), 1),
  Year_Segment = round(vapply(infill_levels, function(p) year_at_logistic(seg_log, p), numeric(1)), 1)
) %>%
  mutate(
    Lag_Years = round(Year_Segment - Year_WholeMile, 1),
    Interpretation = paste0("~", Lag_Years, " yr to fill 0.1-mi gaps after whole miles hit ", Occupancy)
  )

# Model fit summary text (emitted after tables)
fit_summary_md <- paste0(
  "\n##### Model Fit Summary\n\n",
  sprintf("- **Segment linear**: +%.1f percentage points/year → 100%% in ~%.0f\n",
          seg_lin$b, seg_lin$year_100),
  sprintf("- **Mile linear (full series)**: +%.1f percentage points/year → 100%% in ~%.0f\n",
          mile_lin$b, mile_lin$year_100),
  sprintf("- **Mile recent trajectory (last 2 yr)**: +%.1f percentage points/year → 100%% in ~%.0f\n",
          mile_recent_rate, mile_recent_year_100),
  sprintf("- **Segment logistic**: midpoint (50%%) ~%.0f; 95%% ~%.0f\n",
          seg_log$year_50, seg_log$year_95),
  sprintf("- **Mile logistic**: midpoint (50%%) ~%.0f; 95%% ~%.0f\n",
          mile_log$year_50, mile_log$year_95),
  sprintf("\n**Current cumulative occupancy (%.0f)**: %.1f%% of 0.1-mi segments; %.1f%% of whole river miles.\n",
          last_obs,
          cum_occupancy$Seg_Pct[cum_occupancy$Year == last_obs],
          cum_occupancy$Mile_Pct[cum_occupancy$Year == last_obs])
)
```

#### Forecast to Corridor Saturation

Models fit to cumulative occupancy and projected forward:

- **Linear** — constant annual discovery rate; reaches 100% in finite time.
- **Asymptotic** — approaches 100% with diminishing returns (classic discovery / sampling plateau).
- **Logistic** — S-curve to a 100% carrying capacity; slows as unoccupied corridor shrinks.
- **Recent trajectory (whole miles)** — slope from the last two survey years only; tests the hypothesis that pepperweed is already more ubiquitous than the full-series fit implies, and limited annual effort is still revealing occupied miles quickly.

Pepperweed is already common on nearby ditches, streams, and upper-river reaches; these forecasts treat the LORP corridor as filling toward full known occupancy under continued survey and/or spread.

```{r occupancy-forecast-tables}
#| message: false
#| warning: false
#| results: 'asis'

print(kable(milestones,
            col.names = c("Resolution", "Occupancy Milestone", "Linear", "Asymptotic", "Logistic"),
            align = c("l", "l", "c", "c", "c"),
            caption = "Projected calendar year when cumulative occupancy reaches each milestone (fit to 2018–2025 discovery)"))
```

##### Infill lag (whole mile → 0.1-mi segment)

The **horizontal gap** between the blue (whole-mile) and red (0.1-mi) logistic curves is the years of additional discovery needed to occupy every subsegment once whole-mile occupancy has already reached that level.

```{r occupancy-forecast-infill}
#| message: false
#| warning: false
#| results: 'asis'

print(kable(infill_gaps,
            col.names = c("Occupancy level", "Whole-mile year", "0.1-mi year", "Lag (yr)", "Interpretation"),
            align = c("c", "c", "c", "c", "l"),
            caption = "Time lag between whole-mile and subsegment occupancy (logistic forecasts)"))

cat(fit_summary_md)
```

```{r occupancy-forecast-plot}
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 7
#| fig-alt: "Line chart of cumulative pepperweed occupancy over time for whole river miles and 0.1-mile segments, with linear, logistic, and recent-trajectory forecasts toward 100 percent known occupancy."
plot_df_obs <- cum_occupancy %>%
  select(Year, Seg_Pct, Mile_Pct) %>%
  pivot_longer(-Year, names_to = "Series", values_to = "Pct") %>%
  mutate(Series = recode(Series, Seg_Pct = "0.1-mi segments (observed)", Mile_Pct = "Whole miles (observed)"),
         Type = "Observed")

make_forecast_df <- function(years, pred_fn, series_name) {
  data.frame(Year = years, Pct = pred_fn(years), Series = series_name, Type = "Forecast")
}

# Recent scenario only from last observed year forward (avoid rewriting history)
recent_years <- seq(last_obs, max(horizon_end, ceiling(mile_recent_year_100) + 1), by = 1)

plot_df_fc <- bind_rows(
  make_forecast_df(forecast_years, seg_lin$predict, "0.1-mi segments — linear"),
  make_forecast_df(forecast_years, seg_log$predict, "0.1-mi segments — logistic"),
  make_forecast_df(forecast_years, mile_lin$predict, "Whole miles — linear"),
  make_forecast_df(forecast_years, mile_log$predict, "Whole miles — logistic"),
  make_forecast_df(recent_years, mile_recent_predict, "Whole miles — recent trajectory")
)

# Annotation segments for infill lag at selected occupancy levels
annot_levels <- c(50, 75)
annot_df <- data.frame(
  Pct = annot_levels,
  x_mile = vapply(annot_levels, function(p) year_at_logistic(mile_log, p), numeric(1)),
  x_seg = vapply(annot_levels, function(p) year_at_logistic(seg_log, p), numeric(1))
) %>%
  mutate(
    lag = x_seg - x_mile,
    label = sprintf("+%.0f yr to fill\n0.1-mi gaps", lag),
    x_mid = (x_mile + x_seg) / 2
  )

forecast_plot <- ggplot() +
  geom_hline(yintercept = 100, linetype = "dotted", color = "grey50") +
  geom_hline(yintercept = annot_levels, linetype = "dotted", color = "grey80") +
  # Infill-lag brackets (horizontal distance between logistic curves)
  geom_segment(
    data = annot_df,
    aes(x = x_mile, xend = x_seg, y = Pct, yend = Pct),
    color = "#5d6d7e",
    linewidth = 0.7,
    arrow = arrow(ends = "both", length = unit(0.12, "inches"))
  ) +
  geom_text(
    data = annot_df,
    aes(x = x_mid, y = Pct, label = label),
    vjust = -0.4,
    size = 3.2,
    color = "#34495e",
    fontface = "italic",
    lineheight = 0.95
  ) +
  geom_line(data = plot_df_fc %>% filter(grepl("linear", Series) & !grepl("recent", Series)),
            aes(x = Year, y = Pct, color = Series), linewidth = 0.9, linetype = "dashed") +
  geom_line(data = plot_df_fc %>% filter(grepl("logistic", Series)),
            aes(x = Year, y = Pct, color = Series), linewidth = 0.9, linetype = "solid") +
  geom_line(data = plot_df_fc %>% filter(grepl("recent", Series)),
            aes(x = Year, y = Pct, color = Series), linewidth = 1.15, linetype = "dotdash") +
  geom_point(data = plot_df_obs, aes(x = Year, y = Pct, color = Series), size = 2.5) +
  geom_line(data = plot_df_obs, aes(x = Year, y = Pct, color = Series), linewidth = 1.1) +
  # Mark recent-scenario 100% year
  {
    if (is.finite(mile_recent_year_100) && mile_recent_year_100 <= horizon_end) {
      list(
        geom_vline(xintercept = mile_recent_year_100, linetype = "dotted", color = "#6c3483", alpha = 0.7),
        annotate(
          "text",
          x = mile_recent_year_100,
          y = 8,
          label = sprintf("Recent traj.\n→ 100%% ~%.0f", mile_recent_year_100),
          hjust = -0.05,
          size = 3,
          color = "#6c3483"
        )
      )
    } else {
      list()
    }
  } +
  scale_y_continuous(limits = c(0, 105), breaks = seq(0, 100, 25),
                     labels = function(x) paste0(x, "%")) +
  scale_x_continuous(breaks = pretty(forecast_years, n = 8)) +
  scale_color_manual(values = c(
    "0.1-mi segments (observed)" = "#c0392b",
    "Whole miles (observed)" = "#1a5276",
    "0.1-mi segments — linear" = "#e74c3c",
    "0.1-mi segments — logistic" = "#922b21",
    "Whole miles — linear" = "#3498db",
    "Whole miles — logistic" = "#1a5276",
    "Whole miles — recent trajectory" = "#6c3483"
  )) +
  labs(
    title = "Cumulative Pepperweed Occupancy — Observed and Forecast to 100%",
    subtitle = "Solid = logistic; dashed = full-series linear; dot-dash = recent 2-yr whole-mile trajectory\nGrey brackets = years to fill 0.1-mi subsegments after whole miles reach that occupancy",
    x = "Year",
    y = "Cumulative occupancy",
    color = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10),
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 2))

print(forecast_plot)
```

#### Interpretation

`r paste0("- **Whole-mile occupancy** is already high (~", cum_occupancy$Mile_Pct[cum_occupancy$Year == last_obs], "% known), consistent with pepperweed being fairly ubiquitous along the corridor at coarse scale.")`

`r paste0("- **0.1-mile occupancy** remains lower (~", cum_occupancy$Seg_Pct[cum_occupancy$Year == last_obs], "%). The annotated horizontal gaps are the expected years of additional surveying/spread before every subsegment is known occupied at that same occupancy level.")`

`r sprintf("- **Recent-trajectory scenario**: using only %.0f–%.0f whole-mile discovery (+%.1f pp/yr), cumulative mile occupancy reaches 100%% around **%.0f**. This is the “already ubiquitous / under-surveyed” case — if the last two years’ pace continues, known whole-mile occupancy saturates almost immediately relative to the full-series logistic.", min(recent_miles$Year), max(recent_miles$Year), mile_recent_rate, mile_recent_year_100)`

- Full-series linear/logistic forecasts are slower and more consistent with a sampling plateau; the recent trajectory is an upper-bound discovery scenario, not a separate biological invasion model.
- Increased RAS effort in currently empty miles (see Monitoring Recommendations) will accelerate cumulative discovery even without new colonization.

## River Mile Analysis

```{r river-mile-analysis}
#| message: false
#| warning: false
#| fig-width: 16
#| fig-height: 12
#| fig-alt: "Faceted bar charts of pepperweed site counts by whole river mile (0 to 60) for each survey year along the LORP corridor."

# Read river mile shapefile
rivermiles <- st_read('data/LORP_RiverMiles_revised.shp', quiet = TRUE)

# Convert NAME to numeric for proper sorting
rivermiles <- rivermiles %>%
  mutate(RiverMile = as.numeric(NAME))

# Ensure both datasets have the same CRS (WGS84)
rivermiles <- st_transform(rivermiles, st_crs(pepper_data))

# Perform spatial join to find closest river mile for each pepperweed site
pepper_data_with_rivermiles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  select(-LAYER, -sym, -gpxx.Displ, -RiverMile) %>%  # Remove unnecessary columns and duplicate RiverMile
  rename(RiverMile = NAME)

# Create summary data for faceted histogram - bin river miles into whole miles
river_mile_summary <- pepper_data_with_rivermiles %>%
  st_drop_geometry() %>%
  mutate(RiverMile_Binned = floor(as.numeric(RiverMile))) %>%  # Bin to whole miles
  group_by(Year_Display, RiverMile_Binned) %>%
  summarise(Site_Count = n(), .groups = 'drop') %>%
  # Ensure all year-rivermile combinations are represented (extend to mile 60)
  complete(Year_Display = min(Year_Display):max(Year_Display), 
           RiverMile_Binned = 0:60, 
           fill = list(Site_Count = 0))

# Calculate actual reach boundaries from pepperweed data
# Get the minimum river mile for each reach (northern boundary)
pepper_data_with_reaches_and_miles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

reach_boundaries <- pepper_data_with_reaches_and_miles %>%
  st_drop_geometry() %>%
  group_by(Name) %>%
  summarise(
    Min_RiverMile = min(RiverMile, na.rm = TRUE),
    Max_RiverMile = max(RiverMile, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    Reach_Num = as.numeric(gsub("LORP Reach ", "", Name)),
    # Use minimum river mile as the northern boundary (top of reach)
    Label_Position = Min_RiverMile
  ) %>%
  # Sort by minimum river mile to ensure proper ordering (R1 at top, R6 at bottom)
  arrange(Min_RiverMile) %>%
  # Reassign reach numbers based on actual river mile order (1 at top, 6 at bottom)
  mutate(Reach_Num = row_number()) %>%
  # Create proper reach names
  mutate(Name = paste0("LORP Reach ", Reach_Num))

# Display the reach boundaries table
cat("### Reach Boundaries (Northern Boundaries)\n")
cat("**Table showing minimum river mile for each reach (northern boundary):**\n\n")
print(kable(reach_boundaries, 
            col.names = c("LORP Reach", "Min River Mile", "Max River Mile", "Reach Number", "Label Position"),
            align = c("l", "c", "c", "c", "c"),
            caption = "Northern boundaries (minimum river miles) for each LORP reach"))

# Use these actual boundaries for the plot
cat("\n**Using actual reach boundaries from data:**\n")
print(reach_boundaries)

# Create a simple table showing labels and river miles
reach_labels_table <- reach_boundaries %>%
  select(Name, Label_Position) %>%
  arrange(Label_Position)

cat("\n### Reach Labels and River Miles\n")
cat("**Table showing reach labels and their river mile positions:**\n\n")
print(kable(reach_labels_table, 
            col.names = c("Reach Label", "River Mile"),
            align = c("l", "c"),
            caption = "Reach labels and their river mile positions"))

# Create a simple table from pepper_data grouped by reach
pepper_data_with_miles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

reach_min_miles <- pepper_data_with_miles %>%
  st_drop_geometry() %>%
  group_by(Name) %>%
  summarise(Min_RiverMile = min(RiverMile, na.rm = TRUE), .groups = 'drop') %>%
  arrange(Min_RiverMile)

cat("\n### Reach Minimum River Miles\n")
cat("**Table showing minimum river mile for each reach:**\n\n")
print(kable(reach_min_miles, 
            col.names = c("Reach", "Min River Mile"),
            align = c("l", "c"),
            caption = "Minimum river mile for each reach"))

# Create faceted histogram
if(nrow(river_mile_summary) > 0) {
  # Calculate consistent x-axis limits across all years
  max_sites <- max(river_mile_summary$Site_Count, na.rm = TRUE)
  x_limit <- ceiling(max_sites * 1.1)  # Add 10% padding
  
  # Cap sites at 20 for better small site visibility and reverse y-axis (0 at top)
  river_mile_summary_capped <- river_mile_summary %>%
    mutate(Site_Count_Capped = pmin(Site_Count, 20),
           RiverMile_Binned = factor(RiverMile_Binned, levels = sort(unique(RiverMile_Binned), decreasing = TRUE)))
  
  # Manually set reach boundaries based on the table values
  # LORP Reach 1: 0.1 -> bin 0
  # LORP Reach 2: 3.7 -> bin 3 or 4  
  # LORP Reach 3: 19.7 -> bin 19 or 20
  # LORP Reach 4: 34.5 -> bin 34 or 35
  # LORP Reach 5: 39.0 -> bin 39
  # LORP Reach 6: 49.7 -> bin 49 or 50
  
  reach_boundaries_plot <- data.frame(
    Name = c("Reach 1", "Reach 2", "Reach 3", "Reach 4", "Reach 5", "Reach 6"),
    Closest_Bin = c("0", "4", "20", "35", "39", "50")
  )
  
  cat("\n### Manual Reach Boundaries for Plot\n")
  print(reach_boundaries_plot)
  
  river_plot <- ggplot(river_mile_summary_capped, aes(x = Site_Count_Capped, y = factor(RiverMile_Binned))) +
    geom_col(fill = "red", alpha = 0.8, width = 0.7) +
    # Add horizontal lines for reach boundaries using closest binned values
    geom_hline(data = reach_boundaries_plot, aes(yintercept = Closest_Bin), color = "blue", linetype = "solid", alpha = 0.8, size = 1) +
    # Add reach labels - one per reach, positioned below the lines and well within visible area
    geom_text(data = reach_boundaries_plot %>% 
                mutate(Year_Display = 2018),  # Only show labels on 2018 panel
              aes(x = 10, y = Closest_Bin, label = Name), 
              hjust = 0, vjust = 1.5, color = "blue", size = 6, fontface = "bold") +
    facet_wrap(~ Year_Display, ncol = 8, scales = "fixed") +  # Single row, fixed scales
    scale_x_continuous(limits = c(0, 20), breaks = c(0, 5, 10, 15, 20)) +  # Cap at 20
    scale_y_discrete(breaks = function(x) x[seq(1, length(x), by = 5)]) +  # Every 5 miles
    labs(
      title = "Pepperweed Sites by River Mile and Year",
      subtitle = "Distribution of sites along the LORP river corridor (binned by mile, capped at 20 sites)",
      x = "Number of Sites (capped at 20)",
      y = "River Mile"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 22, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 18, color = "#666666"),
      axis.title = element_text(size = 16, face = "bold"),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 14),
      strip.text = element_text(size = 16, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(25, 25, 25, 25)  # Add margin for better spacing
    )
  
  print(river_plot)
  
  # Summary statistics
  cat("### River Mile Distribution Summary\n")
  cat("**Total River Miles Analyzed**:", length(unique(river_mile_summary$RiverMile_Binned)), "\n")
  cat("**Years Covered**:", min(river_mile_summary$Year_Display), "-", max(river_mile_summary$Year_Display), "\n")
  cat("**River Mile Range**:", min(river_mile_summary$RiverMile_Binned), "-", max(river_mile_summary$RiverMile_Binned), "\n")
  
  # Most active river miles
  top_river_miles <- river_mile_summary %>%
    group_by(RiverMile_Binned) %>%
    summarise(Total_Sites = sum(Site_Count), .groups = 'drop') %>%
    arrange(desc(Total_Sites)) %>%
    head(10)
  
  cat("\n### Top 10 Most Active River Miles\n")
  for(i in 1:nrow(top_river_miles)) {
    cat("- **River Mile", top_river_miles$RiverMile_Binned[i], "**:", top_river_miles$Total_Sites[i], "sites\n")
  }
} else {
  cat("No river mile data available for analysis.\n")
}
```

## 2026 Monitoring Recommendations

The [FY 2026–2027 LORP work plan](https://inyowater.org/wp-content/uploads/2026/06/2026-2027-LORP-Work-Plan-and-Budget_draft_062226.pdf) directs summer RAS (rapid assessment survey) monitoring to focus on areas **without known pepperweed populations**. Study-design crosswalk: [`lorp-pia-completion` Doc 11](https://github.com/inyo-gov/lorp-pia-completion/blob/main/docs/11_fy2627_approved_workplan_reference.md). The tables below are derived from `r params$start_year`–`r params$current_year` occurrence data binned to whole river miles (see [River Mile Analysis](#river-mile-analysis)).

### Priority Survey Areas

River-mile segments with **no pepperweed detections on the river corridor** in any survey year (`r params$start_year`–`r params$current_year`). Only sites within 250 m of a river-mile marker count toward occupancy (off-channel ditch/field points are excluded so they do not falsely mark a mile as occupied). These align with the workplan objective to survey unoccupied corridor.

```{r monitoring-recommendations}
#| message: false
#| warning: false
#| results: 'asis'

# Ensure river-mile assignments exist
if (!exists("pepper_data_with_rivermiles")) {
  if (!exists("rivermiles")) {
    rivermiles <- st_read("data/LORP_RiverMiles_revised.shp", quiet = TRUE) %>%
      mutate(RiverMile = as.numeric(NAME)) %>%
      st_transform(st_crs(pepper_data))
  }
  pepper_data_with_rivermiles <- pepper_data %>%
    st_join(rivermiles %>% select(NAME), join = st_nearest_feature) %>%
    rename(RiverMile = NAME)
}

if (!exists("river_reaches")) {
  reach_url <- "https://inyocounty.maps.arcgis.com/sharing/rest/content/items/90e5870bd5914a928bd97b023f07b807/data"
  river_reaches <- st_read(reach_url, quiet = TRUE) %>%
    filter(Name %in% paste0("LORP Reach ", 1:6)) %>%
    st_transform(st_crs(pepper_data))
}

mile_range <- 0:60
corridor_max_m <- 250  # exclude off-channel sites snapped to a river mile from far away

# Distance to nearest river-mile marker (meters)
rm_utm <- st_transform(rivermiles, 32611)
pepper_utm <- st_transform(pepper_data_with_rivermiles, 32611)
nn_idx <- st_nearest_feature(pepper_utm, rm_utm)
corridor_dist_m <- as.numeric(st_distance(pepper_utm, rm_utm[nn_idx, ], by_element = TRUE))

pepper_corridor <- pepper_data_with_rivermiles %>%
  mutate(
    RiverMile_Binned = floor(as.numeric(RiverMile)),
    dist_m = corridor_dist_m
  ) %>%
  filter(dist_m <= corridor_max_m)

detected_miles <- pepper_corridor %>%
  st_drop_geometry() %>%
  distinct(RiverMile_Binned) %>%
  pull(RiverMile_Binned)

never_detected_miles <- setdiff(mile_range, detected_miles)

group_contiguous_miles <- function(miles) {
  if (length(miles) == 0) {
    return(data.frame())
  }
  miles <- sort(miles)
  starts <- miles[c(1, which(diff(miles) > 1) + 1)]
  ends <- miles[c(which(diff(miles) > 1), length(miles))]
  data.frame(RM_Start = starts, RM_End = ends)
}

survey_segments <- group_contiguous_miles(never_detected_miles)

# Assign reach from mile-marker polygons
mile_reach_lookup <- rivermiles %>%
  st_join(river_reaches %>% select(Name), join = st_nearest_feature) %>%
  st_drop_geometry() %>%
  select(RiverMile, Name)

get_reach_for_mile <- function(mile) {
  idx <- which.min(abs(mile_reach_lookup$RiverMile - mile))
  gsub("LORP Reach ", "R", mile_reach_lookup$Name[idx])
}

if (nrow(survey_segments) > 0) {
  survey_segments <- survey_segments %>%
    mutate(
      RM_Range = ifelse(RM_Start == RM_End,
                        paste0("RM ", RM_Start),
                        paste0("RM ", RM_Start, "–", RM_End)),
      Length_Miles = RM_End - RM_Start + 1,
      Reach = vapply((RM_Start + RM_End) / 2, get_reach_for_mile, character(1)),
      Priority = case_when(
        Length_Miles >= 5 ~ "High",
        Length_Miles >= 2 ~ "Medium",
        TRUE ~ "Low"
      ),
      Survey_Note = "No known populations (all years)",
      Bank_Guidance = "Survey both banks"
    ) %>%
    arrange(desc(Length_Miles), RM_Start)

  survey_table <- survey_segments %>%
    select(Priority, RM_Range, Length_Miles, Reach, Survey_Note, Bank_Guidance)

  print(kable(survey_table,
              col.names = c("Priority", "River Mile Range", "Length (mi)", "Reach", "Population Status", "Bank Guidance"),
              align = c("l", "l", "c", "c", "l", "l"),
              caption = paste0("Recommended 2026 RAS survey segments without known pepperweed (n = ",
                               length(never_detected_miles), " river miles, ",
                               round(100 * length(never_detected_miles) / length(mile_range), 1), "% of 0–60 RM corridor)")))
} else {
  cat("All river miles in the 0–60 corridor have at least one historical detection.\n")
}
```

`r if (exists("never_detected_miles") && length(never_detected_miles) > 0) paste0("**Summary:** ", length(never_detected_miles), " of ", length(mile_range), " river miles (", round(100 * length(never_detected_miles) / length(mile_range), 1), "%) have no recorded pepperweed on the corridor (within 250 m of a river-mile marker).")`

### East / West Bank Planning Notes

Historical surveys do not record bank side in the feature service. For planning purposes, site coordinates were classified as **east** or **west** of the river centerline using the local channel direction at each river mile.

- **Unoccupied miles** (table above): survey **both banks** — absence of detections does not indicate which bank was visited.
- **Partially occupied miles** (table below): where all detections fall on one bank, the opposite bank may warrant targeted survey even though populations are known nearby.

```{r monitoring-one-bank}
#| message: false
#| warning: false
#| results: 'asis'

rivermiles_xy <- rivermiles %>%
  mutate(
    lon = st_coordinates(geometry)[, 1],
    lat = st_coordinates(geometry)[, 2]
  ) %>%
  st_drop_geometry() %>%
  arrange(RiverMile)

get_bank <- function(site_lon, site_lat, mile) {
  idx <- which(rivermiles_xy$RiverMile == mile)
  if (length(idx) == 0) {
    return(NA_character_)
  }
  idx <- idx[1]
  if (idx < nrow(rivermiles_xy)) {
    dx <- rivermiles_xy$lon[idx + 1] - rivermiles_xy$lon[idx]
    dy <- rivermiles_xy$lat[idx + 1] - rivermiles_xy$lat[idx]
  } else {
    dx <- rivermiles_xy$lon[idx] - rivermiles_xy$lon[idx - 1]
    dy <- rivermiles_xy$lat[idx] - rivermiles_xy$lat[idx - 1]
  }
  vx <- site_lon - rivermiles_xy$lon[idx]
  vy <- site_lat - rivermiles_xy$lat[idx]
  cross <- dx * vy - dy * vx
  ifelse(cross > 0, "East", "West")
}

site_coords <- st_coordinates(pepper_corridor)
bank_summary <- pepper_corridor %>%
  st_drop_geometry() %>%
  mutate(
    site_lon = site_coords[, 1],
    site_lat = site_coords[, 2],
    Bank = mapply(get_bank, site_lon, site_lat, RiverMile_Binned)
  ) %>%
  group_by(RiverMile_Binned) %>%
  summarise(
    Sites = n(),
    East = sum(Bank == "East", na.rm = TRUE),
    West = sum(Bank == "West", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Banks_Detected = case_when(
      East > 0 & West > 0 ~ "Both",
      East > 0 ~ "East only",
      West > 0 ~ "West only",
      TRUE ~ "Unknown"
    ),
    Opposite_Bank_To_Survey = case_when(
      East > 0 & West == 0 ~ "West",
      West > 0 & East == 0 ~ "East",
      TRUE ~ NA_character_
    ),
    Reach = vapply(RiverMile_Binned, get_reach_for_mile, character(1))
  )

one_bank_gaps <- bank_summary %>%
  filter(!is.na(Opposite_Bank_To_Survey)) %>%
  arrange(RiverMile_Binned) %>%
  mutate(
    RM_Range = paste0("RM ", RiverMile_Binned),
    Planning_Note = paste0("Detections on ", tolower(Banks_Detected), "; consider ", Opposite_Bank_To_Survey, " bank")
  ) %>%
  select(RM_Range, Reach, Sites, Banks_Detected, Opposite_Bank_To_Survey, Planning_Note)

if (nrow(one_bank_gaps) > 0) {
  print(kable(one_bank_gaps,
              col.names = c("River Mile", "Reach", "Sites", "Banks w/ Detections", "Bank to Survey", "Planning Note"),
              align = c("l", "c", "c", "l", "l", "l"),
              caption = "River miles with detections on a single bank only (2018–2025)"))
} else {
  cat("No single-bank detection patterns were identified.\n")
}
```

### Suggested 2026 Survey Sequence

1. **Reach 6 lower corridor (RM 43–48, 51–52, 54–60)** — largest unoccupied segments; highest workplan priority.
2. **Reach 3 gap miles (RM 21–22, 24)** — short unoccupied segments within otherwise occupied reach; efficient to cover during Reach 3 transit.
3. **Single-bank follow-up** — optional targeted passes on the undetected bank at miles listed above, if crew capacity allows.

### Access and Route Planning (next step)

The current navigation polygons follow the **mainstem river-mile centerline** (±100 m). That is enough to flag *where* to look, but not *how* crews actually move:

- **Secondary channels / side channels** often carry the walkable or boat-accessible path and may hold pepperweed even when the mapped mainstem mile looks empty.
- **Prior RAS tracklogs** (Field Maps / Collector GPS tracks) show which banks and braids were reachable in past years — the practical route network for planning day sheets.

**Recommended workflow once those layers are available:**

1. Overlay secondary-channel lines + historical tracklogs on the 2026 gap polygons.
2. Clip or redraw survey units to follow accessible routes (not axis-aligned mainstem boxes).
3. Tag each unit with `Access` (mainstem / secondary / tracklog-only) and preferred entry points.
4. Re-export the Field Maps layer so staff navigate along known paths through negative space.

*Drop GPX/GeoJSON tracklogs or an AGOL item URL into `data/` (or share the service link) and this export can be rebuilt on those routes.*

### Field Maps Navigation Layer
**Area without RAS detections** = June 2023 [high-water](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/highwaterline_north/FeatureServer) floodplain ([south](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/highwaterline_south/FeatureServer) + north), split by river mile and bank, **minus** bank sides that already have pepperweed detections. The [multithreaded river](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/owens_river_feature/FeatureServer) is included as a reference layer.

```{r survey-nav-export}
#| message: false
#| warning: false
#| results: 'asis'

export_script <- "export_survey_nav.R"
if (file.exists(export_script)) {
  status <- system2("Rscript", export_script, stdout = TRUE, stderr = TRUE)
  dir.create("docs/exports", showWarnings = FALSE, recursive = TRUE)
  for (f in c(
    "ras_2026_survey_nav.geojson",
    "ras_2026_survey_nav_shp.zip",
    "lorp_highwater_floodplain.geojson",
    "owens_river_channels.geojson",
    "ras_2026_banks_with_detections.geojson"
  )) {
    src <- file.path("exports/survey_nav_2026", f)
    if (file.exists(src)) {
      file.copy(src, file.path("docs/exports", f), overwrite = TRUE)
    }
  }
}

nav_path <- "exports/survey_nav_2026/ras_2026_survey_nav.geojson"
if (file.exists(nav_path)) {
  nav <- sf::st_read(nav_path, quiet = TRUE)
  nav_summary <- nav %>%
    sf::st_drop_geometry() %>%
    count(Priority, Bank, name = "Units") %>%
    arrange(match(Priority, c("High", "Medium", "Low")), Bank)
  total_ac <- round(sum(nav$Acres, na.rm = TRUE), 1)

  cat("<div class=\"download-section\">\n")
  cat("<h3>Download — area without RAS detections</h3>\n")
  cat("<p>", nrow(nav), " units (~", total_ac, " acres): high-water floodplain on bank sides with <strong>zero</strong> pepperweed detections for that river mile.</p>\n", sep = "")
  cat("<div class=\"download-links\">\n")
  cat("<a href=\"exports/ras_2026_survey_nav_shp.zip\">Shapefile (zip) — Field Maps</a>\n")
  cat("<a href=\"exports/ras_2026_survey_nav.geojson\">No-detection units GeoJSON</a>\n")
  cat("<a href=\"exports/lorp_highwater_floodplain.geojson\">High-water floodplain</a>\n")
  cat("<a href=\"exports/owens_river_channels.geojson\">River channels</a>\n")
  cat("<a href=\"exports/ras_2026_banks_with_detections.geojson\">Banks with detections</a>\n")
  cat("</div>\n")
  cat("</div>\n\n")

  print(kable(nav_summary,
              col.names = c("Priority", "Bank", "Units"),
              align = c("l", "l", "c"),
              caption = "Area without RAS detections (high water − detected bank side)"))
} else {
  cat("Navigation layer not found. Run `Rscript export_survey_nav.R`.\n")
}
```

## Field Documentation

```{r field-images}
#| message: false
#| warning: false
#| results: 'asis'

# Function to check if a feature has attachments
check_feature_attachments <- function(objectid) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  attachments_url <- paste0(base_url, "/", objectid, "/attachments?f=json")
  
  response <- GET(attachments_url)
  if(response$status_code == 200) {
    attachments_data <- content(response, 'parsed')
    return(length(attachments_data$attachmentInfos) > 0)
  }
  return(FALSE)
}

# Function to get attachments for a feature
get_feature_attachments <- function(objectid) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  attachments_url <- paste0(base_url, "/", objectid, "/attachments?f=json")
  
  response <- GET(attachments_url)
  if(response$status_code == 200) {
    attachments_data <- content(response, 'parsed')
    return(attachments_data$attachmentInfos)
  } else {
    return(NULL)
  }
}

# Function to display an image from attachment
display_attachment_image <- function(objectid, attachment_id, alt_text, width = 500) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  image_url <- paste0(base_url, "/", objectid, "/attachments/", attachment_id)
  alt_esc <- htmltools::htmlEscape(alt_text, attribute = TRUE)

  # Create HTML for image display
  html_content <- paste0(
    '<div style="display: inline-block; margin: 15px; text-align: center;">',
    '<img src="', image_url, '" alt="', alt_esc, '" style="max-width: ', width,
    'px; height: auto; border: 2px solid #ddd; border-radius: 8px;">',
    '</div>'
  )

  # Use cat to output HTML directly
  cat(html_content)
}
```

### Field Documentation Images

Pepperweed observations with field photos:

```{r field-images-display}
#| message: false
#| warning: false
#| results: 'asis'

# Find features with attachments (check more features for up to 40 photos)
features_to_check <- pepper_data %>%
  filter(Year_Display >= params$current_year - 2) %>%  # Include last 2 years
  arrange(desc(Year_Display), desc(Abundance)) %>%
  head(100)  # Check more features to find ones with attachments

features_with_attachments <- c()
for(i in 1:nrow(features_to_check)) {
  objectid <- features_to_check$OBJECTID[i]
  if(check_feature_attachments(objectid)) {
    features_with_attachments <- c(features_with_attachments, objectid)
  }
}

# Display images for features that have attachments (up to 40)
images_displayed <- 0
if(length(features_with_attachments) > 0) {
  for(objectid in features_with_attachments[1:min(40, length(features_with_attachments))]) {
    # Get feature data
    feature <- pepper_data[pepper_data$OBJECTID == objectid,]
    
    # Get attachments
    attachments <- get_feature_attachments(objectid)
    
    if(!is.null(attachments) && length(attachments) > 0) {
      # Get the first image attachment
      first_attachment <- attachments[[1]]

      alt_text <- paste0(
        "Field photo of perennial pepperweed, Feature ID ", objectid,
        ", observed ", format(feature$Date_Display, "%Y-%m-%d"),
        ", abundance ", format_abundance(feature$Abundance)
      )

      # Display image first
      display_attachment_image(objectid, first_attachment$id, alt_text)
      
      # Display metadata below the image
      cat("<div style='text-align: center; margin-bottom: 30px;'>")
      cat("<p><strong>Feature ID:</strong>", objectid, "| <strong>Date:</strong>", format(feature$Date_Display, "%Y-%m-%d"), "| <strong>Abundance:</strong>", format_abundance(feature$Abundance), "</p>")
      
      # Include Notes if available
      if(!is.na(feature$Notes) && feature$Notes != "") {
        cat("<p><strong>Notes:</strong>", feature$Notes, "</p>")
      }
      
      cat("</div>")
      
      images_displayed <- images_displayed + 1
    }
  }
}

if(images_displayed == 0) {
  cat("No field documentation images available for recent observations.\n")
} else {
  cat("\n*Field documentation images provide visual confirmation of pepperweed sites and help verify site characteristics.*\n")
}
```

## Conclusion

The `r params$start_year`-`r params$current_year` data reveals expanding pepperweed distribution within the LORP area. The interactive mapping tool provides land managers with a visualization tool for identifying and prioritizing treatment areas over time.

This report will be updated as new data are added to the live feature service. The data source is available at: [ArcGIS REST Service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0)

### Data Disclaimer

The data presented in this report are collected for monitoring and management purposes by the Inyo County Water Department. While every effort is made to ensure accuracy, field conditions, survey timing, and observer experience may affect data quality. Users should verify critical information independently before making management decisions.

---

**Inyo County Water Department**  
© 2025 Inyo County, California. All rights reserved.

*This analysis is automatically updated when new data is added to the ArcGIS feature service.*
