---
title: "Perennial Pepperweed Mapping"
subtitle: "Noxious Weeds Monitoring 2025-2026 LORP Workplan"
tbl-cap-location: top   
affiliation: "Inyo County Water Department"
date: today
date-modified: today
citation:
  type: report
  container-title: "Data Report"
  publisher: "Inyo County Water Department"
  issued: today
  url: https://github.com/inyo-gov/noxious-weeds
google-scholar: true
params:
  current_year: 2025
  start_year: 2018
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
    css: styles.css
  docx:
    toc: true
    toc-depth: 3
    number-sections: true
    reference-doc: custom-reference.docx
execute:
  echo: false
  warning: false
  message: false
prefer-html: true
---

```{r setup}
#| message: false
#| warning: false
library(sf)
library(dplyr)
library(tidyr)
library(leaflet)
library(leaflet.extras2)
library(lubridate)
library(htmltools)
library(knitr)
library(DT)
library(ggplot2)
library(scales)
library(httr)
```

# Summary

This data report covers the monitoring of **Lepidium latifolium** (perennial pepperweed) in the Lower Owens River Project (LORP) area from `r params$start_year` to `r params$current_year`. The data is sourced from the [ArcGIS Online Noxious Weeds 2025 feature service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0) and provides updated spatial occurrence as of `r params$current_year`. 

> "Perennial pepperweed (Lepidium latifolium), an introduced plant from southeastern Europe and Asia, is invasive throughout the western United States. It can establish in a wide range of environments and is a common problem in flood plains, irrigation structures, pastures, wetlands, riparian areas, roadsides, and residential site. Recent surveys identify perennial pepperweed as a weed problem in nearly all of California, and both the California Department of Food and Agriculture (CDFA) and California Invasive Plant Council (Cal-IPC) list it as a noxious weed of great ecological concern." - [UC IPM](https://ipm.ucanr.edu/home-and-landscape/perennial-pepperweed/#gsc.tab=0)

**California Invasive Plant Council (Cal-IPC) Rating: High**  
High – These species have severe ecological impacts on physical processes, plant and animal communities, and vegetation structure. Their reproductive biology and other attributes are conducive to moderate to high rates of dispersal and establishment. Most are widely distributed ecologically.

**California Department of Food and Agriculture (CDFA) Rating: -***  
* – included in the CCR Section 4500 list of California State Noxious Weeds.

```{r data-loading}
#| message: false
#| warning: false

# ArcGIS Feature Service URL
base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0/query?"

# Query parameters for all data
query_params <- paste0(
  "where=1=1",
  "&outFields=*",
  "&f=geojson",
  "&outSR=4326"
)

# Construct full query URL
query_url <- paste0(base_url, query_params)

# Load data from ArcGIS feature service
pepper_data <- st_read(query_url, quiet = TRUE)

# Process dates - convert Unix timestamps to dates
pepper_data <- pepper_data %>%
  mutate(
    Date_Observed = as.POSIXct(Date/1000, origin='1970-01-01'),
    Date_Created = as.POSIXct(CreationDate/1000, origin='1970-01-01'),
    Year_Observed = year(Date_Observed),
    Year_Created = year(Date_Created),
    # Use creation year as fallback when observation year is missing
    Year_Display = ifelse(is.na(Year_Observed), Year_Created, Year_Observed),
    Date_Display = as.POSIXct(ifelse(is.na(Date_Observed), as.numeric(Date_Created), as.numeric(Date_Observed)), origin='1970-01-01')
  )

# Filter to LORP area based on northernmost current year data point
pepper_current_year_for_bounds <- pepper_data %>%
  filter(Year_Display == params$current_year)

if(nrow(pepper_current_year_for_bounds) > 0) {
  # Get the northernmost latitude from current year data
  northernmost_lat <- max(st_coordinates(pepper_current_year_for_bounds)[,2])
  
  # Filter all data to LORP area (south of northernmost current year point)
  pepper_data <- pepper_data %>%
    filter(st_coordinates(.)[,2] <= northernmost_lat)
  
  cat("LORP Area Filter Applied:\n")
  cat("- Northernmost", params$current_year, "point latitude:", round(northernmost_lat, 4), "°N\n")
  cat("- Filtered dataset now contains", nrow(pepper_data), "observations\n")
} else {
  cat("No", params$current_year, "data found for LORP area filtering\n")
}

# Include all records (no year filtering)
# Separate data by display year (observation year with creation year fallback)
pepper_2018 <- pepper_data %>% filter(Year_Display == 2018)
pepper_2019 <- pepper_data %>% filter(Year_Display == 2019)
pepper_2020 <- pepper_data %>% filter(Year_Display == 2020)
pepper_2021 <- pepper_data %>% filter(Year_Display == 2021)
pepper_2022 <- pepper_data %>% filter(Year_Display == 2022)
pepper_2023 <- pepper_data %>% filter(Year_Display == 2023)
pepper_2024 <- pepper_data %>% filter(Year_Display == 2024)
pepper_2025 <- pepper_data %>% filter(Year_Display == 2025)

# Create recent vs historical comparison
pepper_recent <- pepper_data %>% filter(Year_Display >= 2022)
pepper_historical <- pepper_data %>% filter(Year_Display >= 2018 & Year_Display <= 2021)

# Summary statistics for all years (using display year with fallback)
summary_stats <- pepper_data %>%
  filter(!is.na(Year_Display)) %>%
  group_by(Year_Display) %>%
  summarise(
    Total_Observations = n(),
    Avg_Abundance = round(mean(Abundance, na.rm = TRUE), 1),
    Avg_Height = round(mean(Height, na.rm = TRUE), 1),
    Max_Abundance = max(Abundance, na.rm = TRUE),
    Records_With_Obs_Date = sum(!is.na(Year_Observed)),
    Records_With_Creation_Date_Only = sum(is.na(Year_Observed) & !is.na(Year_Created)),
    .groups = 'drop'
  ) %>%
  arrange(Year_Display)
```

## Data Overview

The dataset includes `r nrow(pepper_data)` total observations across `r length(unique(pepper_data$Year_Display[!is.na(pepper_data$Year_Display)]))` years, with peak survey activity in `r names(sort(table(pepper_data$Year_Display, useNA='ifany'), decreasing=TRUE)[1])` (`r max(table(pepper_data$Year_Display, useNA='ifany'))` records).

## Key Findings Summary

- **Total Observations**: `r nrow(pepper_data)` records across all years (2018-2025)
- **Peak Survey Year**: `r names(sort(table(pepper_data$Year_Display, useNA='ifany'), decreasing=TRUE)[1])` with `r max(table(pepper_data$Year_Display, useNA='ifany'))` observations
- **Recent Activity**: `r nrow(pepper_recent)` observations in 2022-2025
- **Historical Baseline**: `r nrow(pepper_historical)` observations in 2018-2021
- **Data Coverage**: `r round((nrow(pepper_data) - sum(is.na(pepper_data$Year_Display))) / nrow(pepper_data) * 100, 1)`% of records have complete date information (using observation date with creation date fallback)

# Interactive Map Comparison

The following interactive map shows pepperweed populations across all years (`r params$start_year`-`r params$current_year`). Use the layer controls in the top-right corner to toggle between different years. The map includes all `r nrow(pepper_data)` documented observations, with feature creation dates used when observation dates are unavailable.

<div class="download-section">
<h3>Data Downloads</h3>
<div class="download-links">
<a href="pepperweed_data.geojson">📊 Download GeoJSON</a>
<a href="pepperweed_data.kml">🌍 Download KML</a>
</div>
<p>Complete dataset with all `r nrow(pepper_data)` pepperweed observations</p>
</div>

```{r generate-geojson}
#| message: false
#| warning: false
#| echo: false
#| results: hide

# Generate GeoJSON file for download
st_write(pepper_data, "pepperweed_data.geojson", driver = "GeoJSON", delete_dsn = TRUE, quiet = TRUE)

# Generate KML file for Google Earth (will be zipped to KMZ)
st_write(pepper_data, "pepperweed_data.kml", driver = "KML", delete_dsn = TRUE, quiet = TRUE)

# Calculate spatial extent
bbox <- st_bbox(pepper_data)
extent_info <- list(
  north = bbox$ymax,
  south = bbox$ymin, 
  east = bbox$xmax,
  west = bbox$xmin,
  center_lat = (bbox$ymax + bbox$ymin) / 2,
  center_lon = (bbox$xmax + bbox$xmin) / 2
)
```

<div class="spatial-extent">
<h4>Study Area Boundaries</h4>
<ul>
<li><strong>North</strong>: `r round(extent_info$north, 4)`°N</li>
<li><strong>South</strong>: `r round(extent_info$south, 4)`°N</li>
<li><strong>East</strong>: `r round(extent_info$east, 4)`°W</li>
<li><strong>West</strong>: `r round(extent_info$west, 4)`°W</li>
<li><strong>Center</strong>: `r round(extent_info$center_lat, 4)`°N, `r round(extent_info$center_lon, 4)`°W</li>
</ul>
</div>

```{r interactive-map}
#| message: false
#| warning: false

# Add reach information to pepper_data if available
if(exists("pepper_data_with_reaches")) {
  # Merge reach information back to main dataset
  pepper_data <- pepper_data %>%
    left_join(
      pepper_data_with_reaches %>% 
        st_drop_geometry() %>% 
        select(OBJECTID, Name),
      by = "OBJECTID"
    )
}

# Load water quality stations as reference points
wq_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/lorp_wq_view/FeatureServer/0/query?where=1=1&outFields=*&f=geojson&outSR=4326"
wq_stations <- st_read(wq_url, quiet = TRUE)

# Load river reaches for labeling (if not already loaded)
if(!exists("river_reaches")) {
  reach_url <- "https://inyocounty.maps.arcgis.com/sharing/rest/content/items/90e5870bd5914a928bd97b023f07b807/data"
  river_reaches <- st_read(reach_url, quiet = TRUE)
  river_reaches <- st_transform(river_reaches, st_crs(pepper_data))
}

# Get reach centroids for labeling - only LORP reaches 1-6
reach_centroids <- river_reaches %>%
  filter(Name %in% c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6")) %>%
  st_centroid() %>%
  mutate(Reach_Label = gsub("LORP Reach ", "R", Name))

# Create comprehensive map with all years
combined_map <- leaflet(pepper_data) %>%
  # Standard basemaps
  addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery") %>%
  addProviderTiles(providers$OpenStreetMap, group = "OpenStreetMap") %>%
  # Add 2018 markers
  addCircleMarkers(
    data = pepper_2018,
    radius = 5,
    color = "black",
    fillColor = "black",
    fillOpacity = 0.7,
    weight = 2,
      popup = ~paste0(
        "<b>2018 Data</b><br>",
        "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
        "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
        "Species: ", Species, "<br>",
        "Abundance: ", Abundance, "<br>",
        "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
        ifelse(exists("Name") && !is.na(Name), paste0("LORP Reach: ", Name, "<br>"), ""),
        "Notes: ", ifelse(is.na(Notes), "None", Notes)
      ),
    group = "2018 Data"
  ) %>%
  # Add 2019 markers
  addCircleMarkers(
    data = pepper_2019,
    radius = 5,
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2019 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2019 Data"
  ) %>%
  # Add 2020 markers
  addCircleMarkers(
    data = pepper_2020,
    radius = 5,
    color = "red",
    fillColor = "red",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2020 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2020 Data"
  ) %>%
  # Add 2021 markers
  addCircleMarkers(
    data = pepper_2021,
    radius = 5,
    color = "yellow",
    fillColor = "yellow",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2021 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2021 Data"
  ) %>%
  # Add 2022 markers
  addCircleMarkers(
    data = pepper_2022,
    radius = 5,
    color = "orange",
    fillColor = "orange",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2022 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2022 Data"
  ) %>%
  # Add 2023 markers
  addCircleMarkers(
    data = pepper_2023,
    radius = 5,
    color = "green",
    fillColor = "green",
    fillOpacity = 0.7,
    weight = 2,
    popup = ~paste0(
      "<b>2023 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2023 Data"
  ) %>%
  # Add 2024 markers
  addCircleMarkers(
    data = pepper_2024,
    radius = 5,
    color = "blue",
    fillColor = "blue",
    fillOpacity = 0.8,
    weight = 3,
    popup = ~paste0(
      "<b>2024 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2024 Data"
  ) %>%
  # Add 2025 markers
  addCircleMarkers(
    data = pepper_2025,
    radius = 5,
    color = "purple",
    fillColor = "purple",
    fillOpacity = 0.9,
    weight = 3,
    popup = ~paste0(
      "<b>2025 Data</b><br>",
      "Date: ", format(Date_Display, "%Y-%m-%d"), "<br>",
      "Date Source: ", ifelse(is.na(Year_Observed), "Creation Date", "Observation Date"), "<br>",
      "Species: ", Species, "<br>",
      "Abundance: ", Abundance, "<br>",
      "Height: ", ifelse(is.na(Height), "Unknown", paste(Height, "m")), "<br>",
      ifelse(exists("Name") && !is.na(Name), paste0("LORP Reach: ", Name, "<br>"), ""),
      "Notes: ", ifelse(is.na(Notes), "None", Notes)
    ),
    group = "2025 Data"
  ) %>%
  # Add water quality stations as reference points (if data available)
  {if(nrow(wq_stations) > 0) {
    addCircleMarkers(
      data = wq_stations,
      radius = 4,
      color = "white",
      fillColor = "blue",
      fillOpacity = 0.8,
      weight = 2,
      popup = ~paste0("<b>Water Quality Station</b><br>Reference Point"),
      group = "Water Quality Stations"
    )
  } else {
    .  # No stations to add
  }} %>%
  # Add reach labels (centroids)
  addLabelOnlyMarkers(
    data = reach_centroids,
    label = ~Reach_Label,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = "center",
      textOnly = TRUE,
      style = list(
        "color" = "white",
        "font-size" = "14px",
        "font-weight" = "bold",
        "text-shadow" = "1px 1px 2px rgba(0,0,0,0.8)"
      )
    ),
    group = "Reach Labels"
  ) %>%
  addLayersControl(
    baseGroups = c("World Imagery", "OpenStreetMap"),
    overlayGroups = c("2018 Data", "2019 Data", "2020 Data", "2021 Data", "2022 Data", "2023 Data", "2024 Data", "2025 Data", "Reach Labels", if(nrow(wq_stations) > 0) "Water Quality Stations"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomright",
    colors = c("black", "darkred", "red", "yellow", "orange", "green", "blue", "purple", if(nrow(wq_stations) > 0) "blue"),
    labels = c("2018", "2019", "2020", "2021", "2022", "2023", "2024", "2025", if(nrow(wq_stations) > 0) "WQ Stations"),
    title = "Pepperweed Distribution by Year"
  )

combined_map
```

## `r params$current_year` Distribution - plants per site estimated from field observations

The following histogram shows the distribution of abundance values for all pepperweed point locations documented in `r params$current_year`:

```{r abundance-histogram}
#| message: false
#| warning: false

# Filter for current year data only
pepper_current_year <- pepper_data %>%
  filter(Year_Display == params$current_year)

# Create abundance histogram
if(nrow(pepper_current_year) > 0) {
  # Create abundance bins (preserve all bins, exclude empty ones)
  abundance_data <- pepper_current_year %>%
    filter(!is.na(Abundance)) %>%
    mutate(
      Abundance_Bin = case_when(
        Abundance <= 10 ~ "1-10",
        Abundance <= 25 ~ "11-25", 
        Abundance <= 50 ~ "26-50",
        Abundance <= 100 ~ "51-100",
        TRUE ~ ">100"
      ),
      Abundance_Bin = factor(Abundance_Bin,
                            levels = c("1-10", "11-25", "26-50", "51-100", ">100"),
                            ordered = TRUE)
    )
  
  # Only include bins that have data
  abundance_counts <- abundance_data %>%
    count(Abundance_Bin) %>%
    filter(n > 0)
  
  # Create bar chart with only populated bins
  hist_plot <- ggplot(abundance_counts, aes(x = Abundance_Bin, y = n)) +
    geom_bar(stat = "identity", fill = "#2E86AB", color = "white", alpha = 0.8, width = 0.7) +
    labs(
      title = paste("Distribution of Pepperweed Abundance in", params$current_year),
      subtitle = paste("Total sites:", nrow(pepper_current_year)),
      x = "Abundance Range (plants)",
      y = "Number of Sites"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 14, color = "#666666"),
      axis.title = element_text(size = 12),
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank()
    ) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 8))
  
  print(hist_plot)
  
  # Management categories
  small_pops <- sum(pepper_current_year$Abundance <= 100, na.rm = TRUE)
  large_pops <- sum(pepper_current_year$Abundance > 100, na.rm = TRUE)
  
  cat("###", params$current_year, "Site Distribution\n")
  cat("- **Total Sites**:", nrow(pepper_current_year), "\n")
  cat("- **Small Sites (≤ 100 plants)**:", small_pops, "sites (", round(small_pops/nrow(pepper_current_year)*100, 1), "%)\n")
  cat("- **Large Sites (> 100 plants)**:", large_pops, "sites (", round(large_pops/nrow(pepper_current_year)*100, 1), "%)\n")
  
  # Detailed bin counts
  cat("\n### Detailed Abundance Distribution\n")
  for(i in 1:nrow(abundance_counts)) {
    bin_name <- abundance_counts$Abundance_Bin[i]
    bin_count <- abundance_counts$n[i]
    cat("- **", bin_name, " plants**: ", bin_count, " sites\n", sep = "")
  }
  
  # Range information
  cat("\n### Abundance Range\n")
  cat("- **Minimum**:", min(pepper_current_year$Abundance, na.rm = TRUE), "plants\n")
  cat("- **Maximum**:", max(pepper_current_year$Abundance, na.rm = TRUE), "plants\n")
} else {
  cat("No", params$current_year, "data available for abundance analysis.\n")
}
```

The interactive map above provides the primary visualization of pepperweed distribution patterns. 


### Recent Sites
```{r new-sites}
#| message: false
#| warning: false

# Show all recent sites (2024-2025) - prioritize small sites for early intervention
recent_sites <- pepper_data %>%
  filter(Year_Display >= 2024) %>%
  select(Date_Display, Abundance, Height, Notes, Year_Display) %>%
  # Sort by abundance (smallest first) to prioritize satellite sites
  arrange(Abundance, desc(Year_Display))

if(nrow(recent_sites) > 0) {
  datatable(recent_sites,
            caption = "Recent Pepperweed Sites (2024-2025) - Prioritize Small Sites for Early Intervention",
            options = list(pageLength = 15, scrollX = TRUE))
} else {
  cat("No recent sites found.")
}
```




# Data Quality and Limitations

## Data Sources
- **Primary Source**: [ArcGIS Online Noxious Weeds 2025 Feature Service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0)

- **Last Updated**: `r format(as.POSIXct(max(pepper_data$CreationDate, na.rm = TRUE)/1000, origin='1970-01-01'), "%B %d, %Y")`
- **Coordinate System**: WGS84 (EPSG:4326)

## Limitations
- Data represents documented observations only
- Population estimates may vary based on survey timing and conditions
- Some areas may be under-sampled due to access constraints
- Abundance categories are based on field estimates

# River Reach Analysis

```{r river-reach-analysis}
#| message: false
#| warning: false
#| fig-width: 16
#| fig-height: 15

# Read river reach polygons from ArcGIS Online and filter to LORP reaches 1-6
reach_url <- "https://inyocounty.maps.arcgis.com/sharing/rest/content/items/90e5870bd5914a928bd97b023f07b807/data"
river_reaches <- st_read(reach_url, quiet = TRUE)

# Filter to only LORP reaches 1-6 - explicit selection
desired_lorp_reaches <- c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6")
river_reaches <- river_reaches %>%
  filter(Name %in% desired_lorp_reaches)

# Ensure both datasets have the same CRS (WGS84)
river_reaches <- st_transform(river_reaches, st_crs(pepper_data))

# Check if we have reach data
cat("Number of river reaches loaded:", nrow(river_reaches), "\n")
cat("Reach names:", unique(river_reaches$Name), "\n")

# Perform spatial join to determine which reach each pepperweed site is closest to
# Use nearest neighbor join since sites might not fall exactly within reach polygons
pepper_data_with_reaches <- pepper_data %>%
  st_join(river_reaches, join = st_nearest_feature)

# Check how many sites were matched
cat("Sites before filtering:", nrow(pepper_data), "\n")
cat("Sites after spatial join:", nrow(pepper_data_with_reaches), "\n")
cat("Sites with reach assignments:", sum(!is.na(pepper_data_with_reaches$Name)), "\n")

# Filter to only sites within reaches
pepper_data_with_reaches <- pepper_data_with_reaches %>%
  filter(!is.na(Name))  # Only keep sites that fall within a reach

# Create summary data for reach analysis
reach_summary <- pepper_data_with_reaches %>%
  st_drop_geometry() %>%
  group_by(Year_Display, Name) %>%
  summarise(Site_Count = n(), .groups = 'drop') %>%
  # Ensure all year-reach combinations are represented
  complete(Year_Display = unique(Year_Display), 
           Name = unique(Name), 
           fill = list(Site_Count = 0))

# Create faceted histogram for reaches
if(nrow(reach_summary) > 0) {
  # Cap sites at 20 for better small site visibility
  reach_summary_capped <- reach_summary %>%
    mutate(Site_Count_Capped = pmin(Site_Count, 20))
  
  # Create ordered factor for reaches (1 at top, 6 at bottom)
  reach_summary_capped <- reach_summary_capped %>%
    mutate(Name_Ordered = factor(Name, levels = rev(c("LORP Reach 1", "LORP Reach 2", "LORP Reach 3", "LORP Reach 4", "LORP Reach 5", "LORP Reach 6"))))
  
  reach_plot <- ggplot(reach_summary_capped, aes(x = Site_Count_Capped, y = Name_Ordered)) +
    geom_col(fill = "red", alpha = 0.8, width = 0.7) +
    facet_wrap(~ Year_Display, ncol = 8, scales = "fixed") +  # Single row, fixed scales
    scale_x_continuous(limits = c(0, 20), breaks = c(0, 5, 10, 15, 20)) +  # Cap at 20
    scale_y_discrete() +  # Reach names
    labs(
      title = "Pepperweed Sites by LORP Reach and Year",
      subtitle = "Distribution of sites within LORP river reaches (capped at 20 sites)",
      x = "Number of Sites (capped at 20)",
      y = "LORP Reach"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 18, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 16, color = "#666666"),
      axis.title = element_text(size = 14),
      axis.text.y = element_text(size = 11),
      axis.text.x = element_text(size = 11),
      strip.text = element_text(size = 13, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(25, 25, 25, 25)
    )
  
  print(reach_plot)
  
  # Summary statistics
  cat("### LORP Reach Distribution Summary\n")
  cat("**Total Reaches Analyzed**:", length(unique(reach_summary$Name)), "\n")
  cat("**Years Covered**:", min(reach_summary$Year_Display), "-", max(reach_summary$Year_Display), "\n")
  cat("**Reaches with Data**:", length(unique(reach_summary$Name[reach_summary$Site_Count > 0])), "\n")
  
  # Most active reaches
  top_reaches <- reach_summary %>%
    group_by(Name) %>%
    summarise(Total_Sites = sum(Site_Count), .groups = 'drop') %>%
    arrange(desc(Total_Sites))
  
  cat("\n### Reach Activity Summary\n")
  for(i in 1:nrow(top_reaches)) {
    cat("- **", top_reaches$Name[i], "**:", top_reaches$Total_Sites[i], "sites\n")
  }
  
  # Sites not in any reach
  sites_outside_reaches <- nrow(pepper_data) - nrow(pepper_data_with_reaches)
  if(sites_outside_reaches > 0) {
    cat("\n### Sites Outside LORP Reaches\n")
    cat("- **Sites outside reach polygons**:", sites_outside_reaches, "\n")
    cat("- **Percentage of total sites**:", round(sites_outside_reaches/nrow(pepper_data)*100, 1), "%\n")
  }
} else {
  cat("No reach data available for analysis.\n")
}
```

# Occupancy Analysis
```{r occupancy-analysis}
#| message: false
#| warning: false

# Load river miles data if not already loaded
if(!exists("rivermiles")) {
  rivermiles <- st_read('data/LORP_RiverMiles_revised.shp', quiet = TRUE)
  rivermiles <- rivermiles %>%
    mutate(RiverMile = as.numeric(NAME))
  rivermiles <- st_transform(rivermiles, st_crs(pepper_data))
}

# Create 1/10 mile segments for detailed occupancy analysis
# Total of 600 segments across 60 miles (0.1 mile resolution)
segment_size <- 0.1
total_miles <- 60
total_segments <- total_miles / segment_size  # 600 segments

# Create segment boundaries
segment_boundaries <- data.frame(
  Segment_ID = 1:total_segments,
  Mile_Start = seq(0, total_miles - segment_size, by = segment_size),
  Mile_End = seq(segment_size, total_miles, by = segment_size)
)

# Join pepperweed data with river miles and reaches
pepper_data_with_miles_and_reaches <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

# Determine which segment each pepperweed site falls into
pepper_data_with_segments <- pepper_data_with_miles_and_reaches %>%
  st_drop_geometry() %>%
  mutate(
    Segment_ID = floor(RiverMile / segment_size) + 1,
    # Ensure segments are within bounds
    Segment_ID = pmin(Segment_ID, total_segments)
  )

# Overall occupancy by year
overall_occupancy <- pepper_data_with_segments %>%
  group_by(Year_Display) %>%
  summarise(
    Segments_Occupied = n_distinct(Segment_ID),
    Total_Segments = total_segments,
    Occupancy_Percent = round(Segments_Occupied / Total_Segments * 100, 1),
    .groups = 'drop'
  ) %>%
  arrange(Year_Display)

# Reach-specific occupancy
reach_occupancy <- pepper_data_with_segments %>%
  group_by(Year_Display, Name) %>%
  summarise(
    Segments_Occupied = n_distinct(Segment_ID),
    .groups = 'drop'
  ) %>%
  # Calculate total segments per reach
  left_join(
    pepper_data_with_miles_and_reaches %>%
      st_drop_geometry() %>%
      group_by(Name) %>%
      summarise(
        Reach_Mile_Start = min(RiverMile, na.rm = TRUE),
        Reach_Mile_End = max(RiverMile, na.rm = TRUE),
        Reach_Total_Segments = (Reach_Mile_End - Reach_Mile_Start) / segment_size,
        .groups = 'drop'
      ),
    by = "Name"
  ) %>%
  mutate(
    Occupancy_Fraction = round(Segments_Occupied / Reach_Total_Segments, 3),
    Occupancy_Percent = round(Occupancy_Fraction * 100, 1)
  ) %>%
  arrange(Year_Display, Name)

# Display results
cat("### Overall Occupancy Analysis\n")
cat("**Resolution**: 1/10 mile segments (600 total segments across 60 miles)\n\n")

# Overall occupancy table
cat("#### Overall Occupancy by Year\n")
print(kable(overall_occupancy, 
            col.names = c("Year", "Segments Occupied", "Total Segments", "Occupancy %"),
            align = c("c", "c", "c", "c"),
            caption = "Number of 1/10 mile segments occupied by pepperweed"))

# Reach occupancy table
cat("\n#### Reach-Specific Occupancy\n")
reach_occupancy_wide <- reach_occupancy %>%
  select(Year_Display, Name, Occupancy_Percent) %>%
  pivot_wider(names_from = Year_Display, values_from = Occupancy_Percent, names_prefix = "Year_") %>%
  arrange(Name)

print(kable(reach_occupancy_wide, 
            col.names = c("LORP Reach", "2018", "2019", "2020", "2021", "2022", "2023", "2024", "2025"),
            align = c("l", rep("c", 8)),
            caption = "Percentage of each reach occupied by pepperweed (by year)"))

# Summary statistics
cat("\n### Occupancy Summary Statistics\n")
cat("**Average Annual Occupancy**:", round(mean(overall_occupancy$Occupancy_Percent), 1), "%\n")
cat("**Highest Occupancy Year**:", overall_occupancy$Year_Display[which.max(overall_occupancy$Occupancy_Percent)], 
    "(", max(overall_occupancy$Occupancy_Percent), "%)\n")
cat("**Lowest Occupancy Year**:", overall_occupancy$Year_Display[which.min(overall_occupancy$Occupancy_Percent)], 
    "(", min(overall_occupancy$Occupancy_Percent), "%)\n")

# Most/least occupied reaches
reach_summary <- reach_occupancy %>%
  group_by(Name) %>%
  summarise(
    Avg_Occupancy = round(mean(Occupancy_Percent), 1),
    Max_Occupancy = max(Occupancy_Percent),
    .groups = 'drop'
  ) %>%
  arrange(desc(Avg_Occupancy))

cat("\n**Most Occupied Reach**:", reach_summary$Name[1], "(", reach_summary$Avg_Occupancy[1], "% average)\n")
cat("**Least Occupied Reach**:", reach_summary$Name[nrow(reach_summary)], "(", reach_summary$Avg_Occupancy[nrow(reach_summary)], "% average)\n")
```

# River Mile Analysis

```{r river-mile-analysis}
#| message: false
#| warning: false
#| fig-width: 16
#| fig-height: 12

# Read river mile shapefile
rivermiles <- st_read('data/LORP_RiverMiles_revised.shp', quiet = TRUE)

# Convert NAME to numeric for proper sorting
rivermiles <- rivermiles %>%
  mutate(RiverMile = as.numeric(NAME))

# Ensure both datasets have the same CRS (WGS84)
rivermiles <- st_transform(rivermiles, st_crs(pepper_data))

# Perform spatial join to find closest river mile for each pepperweed site
pepper_data_with_rivermiles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  select(-LAYER, -sym, -gpxx.Displ, -RiverMile) %>%  # Remove unnecessary columns and duplicate RiverMile
  rename(RiverMile = NAME)

# Create summary data for faceted histogram - bin river miles into whole miles
river_mile_summary <- pepper_data_with_rivermiles %>%
  st_drop_geometry() %>%
  mutate(RiverMile_Binned = floor(as.numeric(RiverMile))) %>%  # Bin to whole miles
  group_by(Year_Display, RiverMile_Binned) %>%
  summarise(Site_Count = n(), .groups = 'drop') %>%
  # Ensure all year-rivermile combinations are represented (extend to mile 60)
  complete(Year_Display = min(Year_Display):max(Year_Display), 
           RiverMile_Binned = 0:60, 
           fill = list(Site_Count = 0))

# Calculate actual reach boundaries from pepperweed data
# Get the minimum river mile for each reach (northern boundary)
pepper_data_with_reaches_and_miles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

reach_boundaries <- pepper_data_with_reaches_and_miles %>%
  st_drop_geometry() %>%
  group_by(Name) %>%
  summarise(
    Min_RiverMile = min(RiverMile, na.rm = TRUE),
    Max_RiverMile = max(RiverMile, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    Reach_Num = as.numeric(gsub("LORP Reach ", "", Name)),
    # Use minimum river mile as the northern boundary (top of reach)
    Label_Position = Min_RiverMile
  ) %>%
  # Sort by minimum river mile to ensure proper ordering (R1 at top, R6 at bottom)
  arrange(Min_RiverMile) %>%
  # Reassign reach numbers based on actual river mile order (1 at top, 6 at bottom)
  mutate(Reach_Num = row_number()) %>%
  # Create proper reach names
  mutate(Name = paste0("LORP Reach ", Reach_Num))

# Display the reach boundaries table
cat("### Reach Boundaries (Northern Boundaries)\n")
cat("**Table showing minimum river mile for each reach (northern boundary):**\n\n")
print(kable(reach_boundaries, 
            col.names = c("LORP Reach", "Min River Mile", "Max River Mile", "Reach Number", "Label Position"),
            align = c("l", "c", "c", "c", "c"),
            caption = "Northern boundaries (minimum river miles) for each LORP reach"))

# Use these actual boundaries for the plot
cat("\n**Using actual reach boundaries from data:**\n")
print(reach_boundaries)

# Create a simple table showing labels and river miles
reach_labels_table <- reach_boundaries %>%
  select(Name, Label_Position) %>%
  arrange(Label_Position)

cat("\n### Reach Labels and River Miles\n")
cat("**Table showing reach labels and their river mile positions:**\n\n")
print(kable(reach_labels_table, 
            col.names = c("Reach Label", "River Mile"),
            align = c("l", "c"),
            caption = "Reach labels and their river mile positions"))

# Create a simple table from pepper_data grouped by reach
pepper_data_with_miles <- pepper_data %>%
  st_join(rivermiles, join = st_nearest_feature) %>%
  st_join(river_reaches, join = st_nearest_feature) %>%
  mutate(RiverMile = as.numeric(NAME)) %>%
  filter(!is.na(Name))  # Only sites with reach assignments

reach_min_miles <- pepper_data_with_miles %>%
  st_drop_geometry() %>%
  group_by(Name) %>%
  summarise(Min_RiverMile = min(RiverMile, na.rm = TRUE), .groups = 'drop') %>%
  arrange(Min_RiverMile)

cat("\n### Reach Minimum River Miles\n")
cat("**Table showing minimum river mile for each reach:**\n\n")
print(kable(reach_min_miles, 
            col.names = c("Reach", "Min River Mile"),
            align = c("l", "c"),
            caption = "Minimum river mile for each reach"))

# Create faceted histogram
if(nrow(river_mile_summary) > 0) {
  # Calculate consistent x-axis limits across all years
  max_sites <- max(river_mile_summary$Site_Count, na.rm = TRUE)
  x_limit <- ceiling(max_sites * 1.1)  # Add 10% padding
  
  # Cap sites at 20 for better small site visibility and reverse y-axis (0 at top)
  river_mile_summary_capped <- river_mile_summary %>%
    mutate(Site_Count_Capped = pmin(Site_Count, 20),
           RiverMile_Binned = factor(RiverMile_Binned, levels = sort(unique(RiverMile_Binned), decreasing = TRUE)))
  
  # Manually set reach boundaries based on the table values
  # LORP Reach 1: 0.1 -> bin 0
  # LORP Reach 2: 3.7 -> bin 3 or 4  
  # LORP Reach 3: 19.7 -> bin 19 or 20
  # LORP Reach 4: 34.5 -> bin 34 or 35
  # LORP Reach 5: 39.0 -> bin 39
  # LORP Reach 6: 49.7 -> bin 49 or 50
  
  reach_boundaries_plot <- data.frame(
    Name = c("Reach 1", "Reach 2", "Reach 3", "Reach 4", "Reach 5", "Reach 6"),
    Closest_Bin = c("0", "4", "20", "35", "39", "50")
  )
  
  cat("\n### Manual Reach Boundaries for Plot\n")
  print(reach_boundaries_plot)
  
  river_plot <- ggplot(river_mile_summary_capped, aes(x = Site_Count_Capped, y = factor(RiverMile_Binned))) +
    geom_col(fill = "red", alpha = 0.8, width = 0.7) +
    # Add horizontal lines for reach boundaries using closest binned values
    geom_hline(data = reach_boundaries_plot, aes(yintercept = Closest_Bin), color = "blue", linetype = "solid", alpha = 0.8, size = 1) +
    # Add reach labels - one per reach, positioned below the lines and well within visible area
    geom_text(data = reach_boundaries_plot %>% 
                mutate(Year_Display = 2018),  # Only show labels on 2018 panel
              aes(x = 10, y = Closest_Bin, label = Name), 
              hjust = 0, vjust = 1.5, color = "blue", size = 6, fontface = "bold") +
    facet_wrap(~ Year_Display, ncol = 8, scales = "fixed") +  # Single row, fixed scales
    scale_x_continuous(limits = c(0, 20), breaks = c(0, 5, 10, 15, 20)) +  # Cap at 20
    scale_y_discrete(breaks = function(x) x[seq(1, length(x), by = 5)]) +  # Every 5 miles
    labs(
      title = "Pepperweed Sites by River Mile and Year",
      subtitle = "Distribution of sites along the LORP river corridor (binned by mile, capped at 20 sites)",
      x = "Number of Sites (capped at 20)",
      y = "River Mile"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 22, face = "bold", color = "#2E86AB"),
      plot.subtitle = element_text(size = 18, color = "#666666"),
      axis.title = element_text(size = 16, face = "bold"),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 14),
      strip.text = element_text(size = 16, face = "bold"),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(25, 25, 25, 25)  # Add margin for better spacing
    )
  
  print(river_plot)
  
  # Summary statistics
  cat("### River Mile Distribution Summary\n")
  cat("**Total River Miles Analyzed**:", length(unique(river_mile_summary$RiverMile_Binned)), "\n")
  cat("**Years Covered**:", min(river_mile_summary$Year_Display), "-", max(river_mile_summary$Year_Display), "\n")
  cat("**River Mile Range**:", min(river_mile_summary$RiverMile_Binned), "-", max(river_mile_summary$RiverMile_Binned), "\n")
  
  # Most active river miles
  top_river_miles <- river_mile_summary %>%
    group_by(RiverMile_Binned) %>%
    summarise(Total_Sites = sum(Site_Count), .groups = 'drop') %>%
    arrange(desc(Total_Sites)) %>%
    head(10)
  
  cat("\n### Top 10 Most Active River Miles\n")
  for(i in 1:nrow(top_river_miles)) {
    cat("- **River Mile", top_river_miles$RiverMile_Binned[i], "**:", top_river_miles$Total_Sites[i], "sites\n")
  }
} else {
  cat("No river mile data available for analysis.\n")
}
```

# Field Documentation

```{r field-images}
#| message: false
#| warning: false
#| results: 'asis'

# Function to check if a feature has attachments
check_feature_attachments <- function(objectid) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  attachments_url <- paste0(base_url, "/", objectid, "/attachments?f=json")
  
  response <- GET(attachments_url)
  if(response$status_code == 200) {
    attachments_data <- content(response, 'parsed')
    return(length(attachments_data$attachmentInfos) > 0)
  }
  return(FALSE)
}

# Function to get attachments for a feature
get_feature_attachments <- function(objectid) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  attachments_url <- paste0(base_url, "/", objectid, "/attachments?f=json")
  
  response <- GET(attachments_url)
  if(response$status_code == 200) {
    attachments_data <- content(response, 'parsed')
    return(attachments_data$attachmentInfos)
  } else {
    return(NULL)
  }
}

# Function to display an image from attachment
display_attachment_image <- function(objectid, attachment_id, width = 500) {
  base_url <- "https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0"
  image_url <- paste0(base_url, "/", objectid, "/attachments/", attachment_id)
  
  # Create HTML for image display
  html_content <- paste0(
    '<div style="display: inline-block; margin: 15px; text-align: center;">',
    '<img src="', image_url, '" style="max-width: ', width, 'px; height: auto; border: 2px solid #ddd; border-radius: 8px; box-shadow: 0 4px 8px rgba(0,0,0,0.15);">',
    '</div>'
  )
  
  # Use cat to output HTML directly
  cat(html_content)
}

# Find features with attachments (check more features for up to 40 photos)
cat("## Field Documentation Images\n")
cat("Pepperweed observations with field photos:\n\n")

# Check which features have attachments - look at more features to get up to 40 photos
features_to_check <- pepper_data %>%
  filter(Year_Display >= params$current_year - 2) %>%  # Include last 2 years
  arrange(desc(Year_Display), desc(Abundance)) %>%
  head(100)  # Check more features to find ones with attachments

features_with_attachments <- c()
for(i in 1:nrow(features_to_check)) {
  objectid <- features_to_check$OBJECTID[i]
  if(check_feature_attachments(objectid)) {
    features_with_attachments <- c(features_with_attachments, objectid)
  }
}

# Display images for features that have attachments (up to 40)
images_displayed <- 0
if(length(features_with_attachments) > 0) {
  for(objectid in features_with_attachments[1:min(40, length(features_with_attachments))]) {
    # Get feature data
    feature <- pepper_data[pepper_data$OBJECTID == objectid,]
    
    # Get attachments
    attachments <- get_feature_attachments(objectid)
    
    if(!is.null(attachments) && length(attachments) > 0) {
      # Get the first image attachment
      first_attachment <- attachments[[1]]
      
      # Display image first
      display_attachment_image(objectid, first_attachment$id)
      
      # Display metadata below the image
      cat("<div style='text-align: center; margin-bottom: 30px;'>")
      cat("<p><strong>Feature ID:</strong>", objectid, "| <strong>Date:</strong>", format(feature$Date_Display, "%Y-%m-%d"), "| <strong>Abundance:</strong>", feature$Abundance, "plants</p>")
      
      # Include Notes if available
      if(!is.na(feature$Notes) && feature$Notes != "") {
        cat("<p><strong>Notes:</strong>", feature$Notes, "</p>")
      }
      
      cat("</div>")
      
      images_displayed <- images_displayed + 1
    }
  }
}

if(images_displayed == 0) {
  cat("No field documentation images available for recent observations.\n")
} else {
  cat("\n*Field documentation images provide visual confirmation of pepperweed sites and help verify site characteristics.*\n")
}
```

# Conclusion

The `r params$start_year`-`r params$current_year` data reveals expanding pepperweed distribution within the LORP area. The interactive mapping tool provides land managers with a visualization tool for identifying and prioritizing treatment areas over time.

This report will be updated as new data are added to the live feature service. The data source is available at: [ArcGIS REST Service](https://services.arcgis.com/0jRlQ17Qmni5zEMr/arcgis/rest/services/Noxious_Weeds_2025_view/FeatureServer/0)

## Data Disclaimer

The data presented in this report are collected for monitoring and management purposes by the Inyo County Water Department. While every effort is made to ensure accuracy, field conditions, survey timing, and observer experience may affect data quality. Users should verify critical information independently before making management decisions.

---

**Inyo County Water Department**  
© 2025 Inyo County, California. All rights reserved.

*This analysis is automatically updated when new data is added to the ArcGIS feature service.*
